---
title: "R Notebook"
output: html_notebook
---

For now, we are going to create our shapefile dataframe 'shp' by associating the first two Bangladesh hits in the database to the shapefiles for Baliadangi and Rajshashi. THIS IS COMPLETE FICTION, FOR THE PURPOSES OF CODE DEVELOPMENT.

Note the structure of the shapefile dataframe here: first column rowid, second is poly_geom.

```{r}

setwd("C:/Users/0146156/Documents/mb_checks/reuters-news-data_disease_gis")

grab_polygon <-function(geom, id) {
  geom %>% dplyr::select(poly_geom = geom) %>% 
           mutate(point_geom = st_centroid(poly_geom),
                  precision = "centroid",
                  reuters_id = id,
                  country_iso = "BGD")
}

bgd_gadm_2 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_2") %>%
              janitor::clean_names()

bgd_gadm_3 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_3") %>%
              janitor::clean_names()

bgd_6 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Baliadangi")) %>%
  grab_polygon(6) %>%
  dplyr::select(poly_geom) %>%
  mutate(rowid = "4435529") %>%
  dplyr::select(rowid, poly_geom)

bgd_90 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Rajshahi")) %>%
  grab_polygon(90) %>%
  dplyr::select(poly_geom) %>%
  mutate(rowid = "4448154") %>%
  dplyr::select(rowid, poly_geom)

shp_old <- rbind(bgd_6, bgd_90)

rm(bgd_6, bgd_90, bgd_gadm_2, bgd_gadm_3)

shp_old

```

Flag our two fictional poly occurrence as poly = TRUE and create a data frames of polygon occurrences only and point occurrences only. 

```{r}

dat <- hits %>% 
  mutate(poly = ifelse(rowid == "4435529" | rowid == "4448154", TRUE, FALSE)) 

dat_poly <- dat %>% filter(poly) 
dat_pt <- dat %>% filter(!poly)

dat_pt

```


This first time through we will do it with no buffer and see how many pixels are covered.

```{r}

shp <- st_sf(dat_poly$poly_geom, crs = 4326)

pt_list <- lapply(1:nrow(shp),

  function(i, shp, template) {
    poly <- shp[i, ] #  grab one polygon
    
    # rasterize the polygon and get coordinates
    tmp <- raster::rasterize(poly, template)
    pts <- xyFromCell(tmp, which(!is.na(getValues(tmp))))
    
    return (pts)
    
  },

  shp, template)

pixel_count <- lengths(pt_list)/2

area <- st_area(shp)

stats_no_buffer <- cbind(dat_poly$id, pixel_count, area) %>% as_tibble() %>%
  rename(id = V1) %>%
  mutate(pixels = as.integer(pixel_count), sq_km = as.numeric(area)/1000) %>%
  select(-pixel_count, -area) %>%
  arrange(desc(sq_km))
  
write.csv(stats_no_buffer, "stats_no_buffer.csv")

```



Let's start by looking at what boostraps() normally generates by taking five bootstraps of our occurrence set.

The command below generates a 5 x 2 tibble.

$ splits is a list of 5 
  First list, which is itself a list of 4 items
    1. $ data: A data frame of 9859 observations of our 10 variables
    2. $ in_id: An integer from 1 to 9859
    3. $ out_id: logical vector, all NA
    4. $ id: a 1 x 1 tibble that ocntains the word "Bootstrap1"
  Second list, same structure as first
  ....
$ id is a character vector with the words "Bootstrap1", "Bootstrap2" etc.

```{r}

generic_bootstrap <- bootstraps(dat_new, times = 2)

generic_bootstrap %>% str

```

How does a single run of our home grown bootstrap compare? 

It looks to be of identical structure.


```{r}

# this function samples one set of coordinates per occurrence ID 
one_per_occ <- function(data) {
  sample <- data %>% group_by(id) %>% slice_sample(n = 1) %>% ungroup()
  return(sample)
}

one_split <- bootstraps(one_per_occ(dat_new), times = 1, simplify = TRUE)

one_split %>% str

```

generic_bootstrap$splits[[1]]$data


This code maintains the class and runs but produces no metrics or predictions after the first split.


```{r}

# this function takes a single sample, using one set of coordinates per occurrence ID
one_per_occ <- function(data) {
  sample <- data %>% group_by(id) %>% slice_sample(n = 1) %>% ungroup()
  return(sample)
}

n_model = 2
all_boots_old <- NULL

# for each bootstrap we start with a fresh geographic sample
for (i in 1:n_model) {
  one_split <- bootstraps(one_per_occ(combined), times = 1, simplify = TRUE) %>%
    mutate(id = paste("Bootstrap", i, sep = ""))
  all_boots_old <- rbind(all_boots_old, one_split)
}

attributes(all_boots_old)

#attributes(all_boots_old)$class <- c("bootstraps", "rset", "tbl_df", "tbl", "data.frame")
attributes(all_boots_old)$times <- n_model
#attributes(all_boots_old)$apparent = FALSE
#attributes(all_boots_old)$strata = FALSE

attributes(all_boots_old)

bag_res <- fit_resamples(
  bag_wf,
  all_boots_old,
  control = control_resamples(save_pred = TRUE)
)

```

Iniitally this code would not run because the class is not maintained. Now that we have assigned the attributes manually we are getting a new error.


```{r}

# this function takes a single sample, using one set of coordinates per occurrence ID
one_per_occ <- function(data) {
  sample <- data %>% group_by(id) %>% slice_sample(n = 1) %>% ungroup()
  return(sample)
}

n_model = 2
rm(all_boots_new)
all_boots_new <- NULL
#rm(all_boots_new)

# for each bootstrap we start with a fresh geographic sample
for (i in 1:n_model) {
  one_split <- bootstraps(one_per_occ(combined), times = 1, simplify = TRUE) %>%
    mutate(id = paste("Bootstrap", str_pad(i, 2, pad = "0"), sep = ""))
  all_boots_new <- rbind(all_boots_new, one_split)
}

attributes(all_boots_new)

attributes(all_boots_new)$class <- c("bootstraps", "rset", "tbl_df", "tbl", "data.frame")
attributes(all_boots_new)$times <- n_model
attributes(all_boots_new)$apparent = FALSE
attributes(all_boots_new)$strata = FALSE
#attributes(all_boots_new)$fingerprint = "a9c7793bdc669d100b7a9d24a688a877"

all_boots_new

bag_res <- fit_resamples(
  bag_wf,
  all_boots_new,
  control = control_resamples(save_pred = TRUE)
)

```

Let's try breaking up the loop. 

This runs, but once again the metrics and predictions are empty after the first run.

I noticed that the 'times' attribute is wrong here and on someother split objects so fixed it here, to no avail. 

```{r}

all_boots_noloop <- NULL

one_split <- bootstraps(one_per_occ(combined), times = 1, simplify = TRUE)

second_split <- bootstraps(one_per_occ(combined), times = 1, simplify = TRUE) %>% 
  mutate(id = "Bootstrap2")

all_boots_noloop <- rbind(one_split, second_split)

attributes(all_boots_noloop)$times <- 2

bag_res <- fit_resamples(
  bag_wf,
  all_boots_noloop,
  control = control_resamples(save_pred = TRUE)
)

```

Let's try some new approaches. 

```{r}


# this function takes a single sample, using one set of coordinates per occurrence ID
one_per_occ <- function(data) {
  sample <- data %>% group_by(id) %>% slice_sample(n = 1) %>% ungroup()
  return(sample)
}

n_model = 2
all_boots_new <- NULL
#rm(all_boots_new)

# for each bootstrap we start with a fresh geographic sample
for (i in 1:n_model) {
  one_split <- bootstraps(one_per_occ(combined), times = 1, simplify = TRUE) %>%
    mutate(id = paste("Bootstrap", str_pad(i, 2, pad = "0"), sep = ""))
  all_boots_new <- rbind(all_boots_new, one_split)
}

attributes(all_boots_new)$class <- c("bootstraps", "rset", "tbl_df", "tbl", "data.frame")
attributes(all_boots_new)$times <- n_model
attributes(all_boots_new)$apparent = FALSE
attributes(all_boots_new)$strata = FALSE
#attributes(all_boots_new)$fingerprint = "a9c7793bdc669d100b7a9d24a688a877"

all_boots_new

bag_spec <- bag_tree(min_n = 10) %>%
  set_engine("rpart", times = 1) %>%
  set_mode("classification")

bag_wf <- workflow() %>%
  add_recipe(tree_rec) %>%
  add_model(bag_spec)

bag_res <- fit_resamples(
  bag_wf,
  all_boots_new
  control = control_resamples(save_pred = TRUE)
)

```


What do our results like if we use the garden variety bootstrap and the same sample of occurrence locations every time? 

```{r}

one_per_occ <- function(data) {
  sample <- data %>% group_by(id) %>% slice_sample(n = 1) %>% ungroup()
  return(sample)
}

n_model = 2

generic_boots <- bootstraps(one_per_occ(combined), times = n_model, simplify = TRUE)

bag_res_generic <- fit_resamples(
  bag_wf,
  generic_boots,
  control = control_resamples(save_pred = TRUE)
)



```

Here we modify the second $id variable stored within the first column of the split object itself. 
THIS WORKS

```{r}

# this function takes a single sample, using one set of coordinates per occurrence ID
one_per_occ <- function(data) {
  sample <- data %>% group_by(id) %>% slice_sample(n = 1) %>% ungroup()
  return(sample)
}

n_model = 2
all_boots3 <- NULL

set.seed(456)

# for each bootstrap we start with a fresh geographic sample
for (i in 1:n_model) {
  one_split <- bootstraps(one_per_occ(combined), times = 1, simplify = TRUE) %>%
    mutate(id = paste("Bootstrap", i, sep = ""))
  one_split$splits[[1]]$id <- as.tibble(paste("Bootstrap", i, sep = "")) %>% rename(id = value)
  all_boots3 <- rbind(all_boots3, one_split)
}

attributes(all_boots3)$times <- 2

bag_res_fixed <- fit_resamples(
  bag_wf,
  all_boots3,
  control = control_resamples(save_pred = TRUE)
)

```