---
title: "Ebola sites notebook"
output:
  github_document:
    toc: true
---

```{r}
library(tidyverse)
library(sf)
```

sources: https://figshare.com/articles/dataset/Human_index_cases_-_Ebola_virus_disease/3423593?backTo=/collections/eLife_Advances_-_Updates_to_the_zoonotic_niche_map_of_Ebola_virus_disease_in_Africa/3267965


```{r}
mylne_polys <- st_read("./input_data/Ebola/Mylne/09EboladatapaperGISpolygonshapefile/historical_spread_polygons.shp")
mylne_points <- st_read("./input_data/Ebola/Mylne/08EboladatapaperGISpointsshapefile/historical_spread_points.shp") 
```


## Uganda


### uga_89

Excluding because not enough preceding data. 


## Gabon

Load the Gabon GADM.

```{r}
st_layers("./input_data/Gabon/gadm36_GAB_gpkg/gadm36_GAB.gpkg")
```

```{r}
gadm36_GAB_2 <- st_read("./input_data/Gabon/gadm36_GAB_gpkg/gadm36_GAB.gpkg", layer = "gadm36_GAB_2")
```




### gab_86

This is a 2001 outbreak in Gabon. The Metabiota data cites it as Etakangaye. 

The MetaBiota data does not have coordinates, but Shapiro cites Mylne for the location. 

We can reference [Mylne's data here](https://doi.org/10.6084/m9.figshare.1168886). 

From the data notes:

"GE and the digitized article map were used to geo-ref. Because we are not 100% sure where the village is, a polygon needs to be drawn around the excel coords. (1.07168, 14.10658)"

We extract the polygon and add the centroid. 

```{r}
gab_86_poly <- mylne_polys %>%
  filter(uniq_id == 164) %>%
  select(poly_geom = geometry) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid")
```

### gab_85

Another 2001 outbreak in Gabon, this time in Ekata.

From the Mylne notes:

"Digitized from Pourrut 2005." The link to [that paper is here](https://www.sciencedirect.com/science/article/abs/pii/S1286457905001437?via%3Dihub).

Appears to be an exact site. Have asked Allison for Pourrut paper and will update when I know more. 

0.67705	14.28902


```{r}
gab_85_point <- mylne_points %>%
  filter(uniq_id == 162) %>%
  select(point_geom = geometry) %>%
  mutate(precision = "exact")
```


### gab_82

One last 2001 outbreak in Gabon, this time in Memdemba. 

From the Mylne notes: 

"Pourrut 2005 used to locate actual site of index case."

```{r}
gab_82_point <- mylne_points %>%
  filter(uniq_id == 160) %>%
  select(point_geom = geometry) %>%
  mutate(precision = "exact")
```




## Republic of the Congo

### cod_84


This is a 2001 outbreak in the Republic of the Congo. Metabiota identified it as Entsiami, Abolo, Ambomi. 

From the Mylne dataset: 

"Entsiami' digitized from Pourrut 2005 and a 5km radius used to include the neibouring villages of Abolo & Ambomi (0.09141, 14.21818)"

```{r}
cod_84_poly <- mylne_polys %>%
  filter(uniq_id == 161) %>%
  select(poly_geom = geometry) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid")
```


### cod_83

This is another 2001 outbreak in the Republic of the Congo. Metabiota listed it as Oloba, Mbomo district. 

From the Mylne dataset: "Pourrut 2005 used to locate actual site of index case."

```{r}
cod_83_point <- mylne_points %>%
  filter(uniq_id == 163) %>%
  select(point_geom = geometry) %>%
  mutate(precision = "exact")
```




