---
title: "R Notebook"
output: html_notebook

*** test
---

```{r}
library(sf)
library(tidyverse)
```


# Democratic Republic of the Congo

## Health Zones

I  downloaded the DRC health zones data [from the HDX site](https://data.humdata.org/dataset/dr-congo-health-0) on 20/12/2021. 

```{r}
unzip("./input_data/DRC/rdc_zone_de_sante_09092019.zip", exdir = "./input_data/DRC/health_zones") 
```



```{r}
drc_hz_raw <- st_read("./input_data/DRC/health_zones/RDC_Zone_de_sante_09092019.shp") %>%
  janitor::clean_names()
```

```{r}
plot(drc_hz_raw$geometry)
```

From here, we need to get the following shapes:

* Biena Health Zone, North Kivu Province
* Mangina health area, North Kivu province (SEE NOTE BELOW)
* Likati health zone of the province of Bas Uélé
* Isiro health zone †(index)
* Luebo health zone (index)

```{r}
drc_hz_hits <- drc_hz_raw %>%
  filter(nom %in% c("Biena", "Likati", "Isiro", "Luebo"))

```

Just curious: How big are these areas?

```{r}
drc_hz_hits %>%
  mutate(the_area = st_area(.)) %>%
  mutate(the_area_m2 = units::set_units(the_area, "m^2"),
         the_area_km2 = units::set_units(the_area_m2, "km^2")) %>%
  select(nom, the_area_km2)
```
```{r}
drc_hz_hits %>%
  ggplot() + 
  geom_sf()
```




However, it appears Mangina isn't a health area, but rather a town.


Remaining places:
* Mbandaka, Equateur Province
* Bikoro region of Equateur Province
* Mangina, North Kivu province (added from health zones above)
* Ikanamongo Village (Near Boende) in Equateur province
* Bamoukamba 2 (index)

```{r}
unzip("./input_data/DRC/cod_admbnda_rgc_itos_20190911_shp.zip", exdir = "./input_data/DRC/hdx_admin")
```




```{r}
drc_admin_2 <- st_read("./input_data/DRC/hdx_admin/cod_admbnda_adm2_rgc_20190911.shp") %>%
  janitor::clean_names()
```
Remaining places:
* Mbandaka, Equateur Province
* Bikoro region of Equateur Province
* Mangina, North Kivu province (added from health zones above)
* Ikanamongo Village (Near Boende) in Equateur province
* Bamoukamba 2 (index)


```{r}
drc_admin_2 %>%
  filter(adm2_fr %in% c("Mbandaka", "Bikoro"))
```



# Bangladesh

## GADM

```{r}
st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")
```

```{r}
bgd_gadm_3 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg", layer="gadm36_BGD_3") %>%
  janitor::clean_names()
```

```{r}
bgd_gadm_3 %>%
  filter(str_detect(name_3, "Thakurgaon"))
```

Some help here....

https://www.sciencedirect.com/science/article/pii/S1201971220322451#bib0195


```{r}
bgd_41 <- bgd_gadm_3 %>%
  filter(name_3 == "Joypurhat Sadar") %>%
  mutate(rtrs_id = "41")
```








## Villages

Source: http://geoportal.icpac.net/layers/geonode%3Auga_villages_jan_2009

```{r}
unzip("./input_data/Uganda/villages/uga_villages_jan_2009.zip", exdir = "./input_data/Uganda/villages")
```

Now let's import it. 

```{r}
uganda_villages <- st_read("./input_data/Uganda/villages/uga_villages_jan_2009.shp") %>%
  janitor::clean_names()
```

```{r}
uganda_villages %>%
  filter(str_detect(name, "moyo"))
```



