---
title: "Marburg sites notebook"
output:
  github_document:
    toc: true
---

```{r}
library(tidyverse)
library(sf)
library(arcpullr)
```


## Uganda

We will need this GADM layer, so let's go ahead and load it. 

```{r}
uga_gadm_4 <- st_read("./input_data/Uganda/gadm/gadm36_UGA.gpkg", layer="gadm36_UGA_4") %>%
  janitor::clean_names()
```


### uga_x6

Let's start with the 2017 outbreak in Kween. [According to this paper](https://researchonline.lshtm.ac.uk/id/eprint/4653965/1/Marburg-virus-disease-outbreak.pdf):

"The initial probable index case was a 35-year-old male herdsman (cattle and goats)
from Tulwo village, Kamwam Parish, Kaproron Sub-county of Kween District who undertook
frequent hunting forays to the Kaptum grazing grounds in Kween District, which include a
cave inhabited by bats of Rousettus species. The investigation established that he was also a
crop farmer. He lived in an area close to a cave, inhabited by Rousettus species fruit bats, on
the slopes of Mount Elgon. Contacts reported he frequented the interior of the cave with chil-
dren to mine rock salt for livestock and collect bat droppings for manure for his crop farm."

So I was able to track it down to Kamwam parish. I could not identify the exact location of Tulwo village, so there is room for improvement here. 


```{r}
uga_x6 <- uga_gadm_4 %>%
  filter(str_detect(name_4, "Kamwam")) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid",
         the_year = 2017,
         id = "uga_x6")
  
```

### uga_21

Next up is a 2014 outbreak in Kampala. 

https://www.who.int/emergencies/disease-outbreak-news/item/10-october-2014-marburg-en
https://pubmed.ncbi.nlm.nih.gov/28518032/
https://wwwnc.cdc.gov/eid/article/23/6/17-0047_article
https://academic.oup.com/trstmh/article/109/6/366/2461644

In particular, the last link points us to the Mawokota district of Mpigi. 

```{r}
uga_21 <- uga_gadm_4 %>%
  filter(str_detect(name_2, "Mawokota")) %>%
  select(name_2) %>%
  group_by(name_2) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid",
         the_year = 2014,
         id = "uga_21")

```





### uga_40

This is a 2012 outbreak.

https://pubmed.ncbi.nlm.nih.gov/26209681/

The closest I can get is Ibanda district. 

```{r}
uga_40 <- uga_gadm_4 %>%
  filter(str_detect(name_2, "Ibanda")) %>%
  select(name_2) %>% 
  group_by(name_2) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid",
         the_year = 2012,
         id = "uga_40")

```


### uga_53

This is from 2008. 

This should be one of two spillover events linked to Python cave. One is an American, the other involved Dutch travelers. This one is for the Dutch travelers. 

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2815969/


In Nyakarahuka, et. al., they published the locations of Marburg cases but no metadata. So the data are simply lats/longs and a species. 

With some deductive reasoning, we can sort some locations out. 


```{r}
nyakarahuka <- readxl::read_xlsx("./input_data/Marburg/Nyakarahuka/S1.xlsx") %>%
  janitor::clean_names() %>%
  mutate(latitude_subbed = str_replace(latitude, "S", "0")) %>%
  mutate(latitude_clean = as.numeric(latitude_subbed)) %>%
  select(species, latitude = latitude_clean, longitude)
```

```{r}
nyakarahuka_marburg <- nyakarahuka %>%
  filter(species=="Marburg case") %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326)
```


```{r}
hits <- st_join(nyakarahuka_marburg, uga_gadm_4) 
```

There's two hits within Queen Elizabeth NP territory. Did some additional sleuthing and pretty confident these represent the Python Cave. Obviously we want to check it out. 

```{r}
uga_53 <- hits %>%
  filter(str_detect(name_4, "Queen")) %>%
  select(geometry) %>%
  distinct() %>%
  select(point_geom = geometry) %>%
  mutate(precision = "exact",
         the_year = 2008,
         id = "uga_53")
```

### uga_59

This outbreak occured in June 2007, [acccording to this document](https://academic.oup.com/trstmh/article/109/6/366/2461644?login=true#supplementary-data).

Based on this article, it sounds like a company called Simba has the rights for the location of the Kitka mine. 

Uganda has its mining cadastral on a Trimble site, which means it's on an arcgis rest server, which means we can use arcpullr to extract it. After some investigating, I figured out this is the url we want:


```{r}
uganda_active_mining <- arcpullr::get_spatial_layer("https://miningcadastre.minerals.go.ug/arcgis/rest/services/Uganda_Licenses/MapServer/2") %>%
  janitor::clean_names()
```


```{r}
uga_59 <- uganda_active_mining %>%
  filter(str_detect(parties, "Simba")) %>%
  select(poly_geom = geoms) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid",
         the_year = 2007,
         id = "uga_59")

```




### uga_60

This is the first spillover involving Python cave. It occured in 2007. 

We can just repeat the data we extracted earlier from uga_53. 

```{r}
uga_60 <- hits %>%
  filter(str_detect(name_4, "Queen")) %>%
  select(geometry) %>%
  distinct() %>%
  select(point_geom = geometry) %>%
  mutate(precision = "exact",
         the_year = 2007,
         id = "uga_60")
```


## Angola 

We only have one known Marburg event in Angola. 

### ago_70

This occurred in 2004. 

According to [this paper](https://journals.asm.org/doi/10.1128/JVI.00069-06), the index case was never determined. For now, we'll use the centroid of the province where the outbreak occurred. 

Let's get the layers of the Angola GADM data. 

```{r}
st_layers("./input_data/Angola/GADM/gadm36_AGO.gpkg")
```

```{r}
angola_gadm <- st_read("./input_data/Angola/GADM/gadm36_AGO.gpkg", layer = "gadm36_AGO_1") %>%
  janitor::clean_names()
```
```{r}
uige_province <- angola_gadm %>%
  filter(gid_1 == "AGO.17_1")
```


```{r}
ago_70 <- uige_province %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid",
         the_year = 2004,
         id = "ago_70")
```

# Put it all together

Let's combine the individual records. 

```{r}
marburg_points <- bind_rows(ago_70, uga_x6, uga_21, uga_40, uga_53, uga_59, uga_60)
```

This command effectively tells R which geometry column is the operative one.

```{r}
st_geometry(marburg_points) <- "point_geom"
```

Export the data.

```{r}
marburg_points %>%
  st_write("./spatial_outputs/marburg.gpkg", append=FALSE)
```







