---
title: "Nipah sites in Bangladesh"
output:
  github_document:
    toc: true
---

```{r}

library(tidyverse)
library(sf)

```

Set working directory for Allison's Windows laptop

```{r}

setwd("C:/Users/0146156/Documents/mb_checks/reuters-news-data_disease_gis")

```

Get GADM data. 

```{r}

st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")

bgd_gadm_2 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_2") %>%
              janitor::clean_names()

bgd_gadm_3 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_3") %>%
              janitor::clean_names()

bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer="gadm36_BGD_4") %>%
              janitor::clean_names()

```

Setting up function that returns variables based on single GADM polygon.

```{r}

grab_polygon <-function(geom, id) {
  geom %>% select(poly_geom = geom) %>% 
           mutate(point_geom = st_centroid(poly_geom),
                  precision = "centroid",
                  reuters_id = id,
                  country_iso = "BGD")
}



```

### bgd_3

Unspecified number of Nipah cases in Bangladesh, 2021, no other details in data.

ProMedMail points to a short Outbreak News piece about an updated Nipah case count for Bangladesh in 2021, listing two cases [here](http://outbreaknewstoday.com/bangladesh-reports-two-nipah-virus-cases-in-2021/)

That is based on this chart from Bangladesh's Institute of Epidemiology, Disease Control and Research, [here](https://iedcr.gov.bd/surveillances/93c87e70-9c22-4f21-9506-a1161ecf404f) 

No way to pinpoint potential spillover location based on this source. 

@waiting AM emailed IEDCR director about location of cases from their statistics, Jan 11
*** For now, leaving this data point out 

### bgd_5

Unspecified number of Nipah cases in Bangladesh in 2020, no other details.

Likely the same source as above but unclear. 

Searched ProMed

@waiting IEDCR director
*** For now, leaving this data point out 

### bgd_6

Unspecified number of Nipah cases in Thakurgaon District in 2019.

Academic paper describes this outbreak, with cases from Jan 15 to March 17 [here](https://www.ijidonline.com/action/showPdf?pii=S1201-9712%2820%2932414-0)

Primary case: "He was a traditional healer, had no history of taking RDPS, occasionally visited forest to collect herbs and animals."

No additional information about location here. 

Mention in this local press report of Baliadangi Upazila Health Complex, in relation to March cases, but paper also notes "In February, five members of a family died mysteriously in Baliadanga upazila" [here](https://www.newagebd.net/article/67534/3-of-a-family-suspected-to-be-infected-with-nipah-virus)

Seems likely that IEDCR can also sort this out. 

@waiting IEDCR director
*** Coding to Baliadangi Upazila

```{r}

bgd_6 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Baliadangi")) %>%
  grab_polygon(6)

```

### bgd_14 to bgd_19

These are all clearly based on an Outbreak News summary of cases in 2015 reported by IEDCR by district [here](http://outbreaknewstoday.com/bangladesh-reports-nine-nipah-virus-cases-to-date-in-2015/)

Nothing further on ProMedMail.

Local media paces a Naogaon cluster (bgd_17) in "Valain village under Manda Upazila" [here](https://www.risingbd.com/english/Nipah_virus_kills_3_in_Naogaon/22046)

Daily Star (article)[https://www.thedailystar.net/2-die-of-nipah-virus-in-naogaon-62907]

May 18, 2015: "Two people died of Nipah virus at Manda upazila in Naogaon last week, said the Institute of Epidemiology, Disease Control and Research today."

@waiting IEDCR director
*** Coding bgd_17 to Manda Upazila, the rest at the district level

```{r}

bgd_14 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Faridpur")) %>%
  grab_polygon(14)

bgd_15 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Rajbari")) %>%
 bgd_17 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Manda")) grab_polygon(15)

bgd_16 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Magura")) %>%
  grab_polygon(16)

bgd_18 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Nilphamari")) %>%
  grab_polygon(18)

bgd_19 <- bgd_gadm_2 %>%
  filter(str_detect(name_2, "Panchagarh")) %>%
  grab_polygon(19)


```

#bgd_24

Data says Dhaka, 2013, no other details.

Local press article from early 2013 notes first ever Nipah death in Dhaka, which is outside the Nipah belt, with this detail: "'Although the sap was brought from Mymensingh [District], it was consumed in Dhaka. This is the first case of a Nipah-related death in the city,' said Alamgir."
[source](https://reliefweb.int/report/bangladesh/fresh-outbreak-nipah-virus-bangladesh)

Earlier report from January: "According to its Director, the family drank raw date sap brought from Bhaluka on Jan 11 and fell ill six days later" - goes on to not ethat one person in Bhaluka supplied 100 bottles to families in Dhaka.

Bhaluka is an upazila within the Mymensingh district, but Bhaluka is also a village in the upazila. 

*** Coding to the upazila

```{r}

bgd_24 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Bhaluka")) %>%
  grab_polygon(19)

```

#bgd_25

Data notes district Faridpur and 2013, no other details.

Nothing on ProMED or online that I can find.

Faridpur is not mentioned in EICDR's last 2013 report, though it was from May 2013 so possibly [incomplete](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11)

*** Not coding yet

#bgd_26

Data notes district Madaripur and 2013, no other details.

Nothing on ProMED or online that I can find.

Faridpur is not mentioned in EICDR's last 2013 report, though it was from May 2013 so possibly [incomplete](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11)

*** Not coding

#bgd_27

Data sais "Makikgank" which is not a district, but district Manikganj did have a documented outbreak that year:

"At Hafeznagar village in Shibalaya upazila of the district, 4 people, including a couple, died of an unknown disease in the last 2 weeks. The victims are a 40-year-old woman, her husband age 45 years, the woman's mother of 55 years, and a neighbour, a man 36-years-old. The neighbor and the woman died at Dhaka Community Hospital on 28 and 29 Mar [2013]. The mother died on way to Dhaka on 28 Mar [2013], while the husband died at Monno General Hospital in Manikganj on 14 Mar [2013], family sources said." [source](https://promedmail.org/promed-post/?id=20130401.1613261)

Follow up post confirms this was Nipah, see [here](https://promedmail.org/promed-post/?id=1615535)

Hafeznagara is not in Google Maps or the [Manikganj small area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Manikganj.pdf)

*** Coding to Shibalaya upazila

```{r}

bgd_27 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Shibalaya")) %>%
  grab_polygon(27)

```

#bgd_28

Data notes district Shariatpur and 2013, no other details.

Nothing in ProMED, Daily Star, general Google search, Factiva, IECDR report from May 2013.

*** Not coding 

#bgd_29 

Data notes district Kushtia and 2013, no other details.

*** Coding to Kushtia district

```{r}

bgd_29 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Kushtia")) %>%
  grab_polygon(29)

```
#bgd_30

Magura district, 2013, no other details.

One case is described in ProMED:
"Nipah virus claimed the life of a minor boy at Ghosgati village in Shalikha upazila on Sunday [10 Mar 2013]. A boy, aged 2, of the village died at Jessore General Hospital in the morning. District Civil Surgeon Dr Muhammad Shafiqul Alam said the boy had fallen sick after drinking date [palm] juice on Thursday [7 Mar 2013]. He was admitted to Magura district hospital."
[source](https://promedmail.org/promed-post/?id=1592574)

Magura only appears once on the IECDR list for the year [here](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11) so this is likely the right case

The small area atlas lists a "Ghoshgati" in Gangarampur Union of Shalikha upazila. Mindat lists coordinates for a Ghoshgati in Khulna but that's in the wrong upazila. 

*** Coding to Gangarampur Union

```{r}

bgd_30 <- bgd_gadm_4 %>% filter(str_detect(name_2, "Magura")) %>%
  filter(str_detect(name_4, "Gangarampur")) %>%
  grab_polygon(30)

```
#bgd_31

Chapainawabganj district, 2013 no other details.

Alternate transliteration is Chapai Nawabganj.

The district is not on IECDR's 2013 case list but it is on the 2014 one, which actually covers December 2013 as well: [here](http://www.iedcr.org/pdf/files/nipah/Nipah%20situation_18.02.2014.pdf)

Nothing in the Daily Star, Factiva, ProMED.

*** Not coding yet

#bgd_32




Full IECDR location list for 2013:
[source](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11)
Rajbari - district
Jhenaidah - district
Naogaon - district
Nator - probably Natore, district
Natore - district
Gaibandha - distict
Mymensingh - city and division
Pabna - city, district
Rajshahi - division, district and city
Nilphamari - district
Kurigram - district
Kushtia - district
Magura - district
Manikganj - district





ProMED-mail:

"Among the casualties, there have been 2 from each of Dhaka, Pabna, and Natore and one from each of Rajbari, Jhenaidah, Naogaon, and Rajshahi, according to Bangladesh's Institute of Epidemiology Diseases Control and Research (IEDCR)."

Pabna (city and district)
Natore (district)
Rajbari (district)
Jhenadidah (district)
Naogaon (district)
Rajshahi (city, district and division)










```

### bgd_87

[This paper identifies an outbreak](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010) in Meherpur, Bangladesh from April-May 2001. 

"The first NiV outbreak was reported in April 2001 from a village in district Meherpur, Bangladesh with 13 confirmed cases and 9 (69.2%) deaths."

What we don't know which Meherpur, though. There's Meherpur at levels 2, 3 and 4. 


```{r}
bgd_gadm_4 %>%
  filter(str_detect(name_3, "Naogaon"))

```
The above paper leads us to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf), which says:

"Surveys took place in the villages of Chandpur (population 604), where persons who died or were hospitalized had resided..."

So that's a clue. 