---
title: "Nipah sites notebook"
output:
  github_document:
    toc: true
---

```{r}
library(tidyverse)
library(sf)
```

What I've tried to do here is first group incidents by country, then do an individual section for each event. You'll see that each section is headlined by the ISO 3-digit country code and the unique identifier I added to the Metabiota dataset in Google Sheets.

For general administrative boundaries, I'm using [the GADM database](https://gadm.org/). When I start a new country, I download that country's geopackage into the directory "./input_data/country/GADM/", where I extract the file to that directory instead of letting the unzipper create a new directory. 

Because these files are large, you'll want to use [Github's Large File Storage](https://git-lfs.github.com/) to track them. I would do the following commands:

* git lfs track "*.zip"
* git lfs track "*.gpkg"

As of now, my goal is to have each incident with the following fields:

* unique identifier
* polygon geometry, if used
* point geometry

We may want to consider adding a memo field to contain important information about each incident but for now I've not done that. 

What you might do is create a new R notebook and just start the document like below (i.e. ignore all the stuff above here). Then when you're done, we can just append it to this doc. 



## India

https://censusindia.gov.in/2011census/findvillages.aspx
https://data.gov.in/catalog/complete-villages-directory-indiastatedistrictsub-district-level-census-2011?filters%5Bfield_catalog_reference%5D=166603&format=json&offset=0&limit=6&sort%5Bcreated%5D=desc

https://stategisportal.nic.in/stategisportal/


First, let’s check the layers.

```{r}
st_layers("./input_data/India/gadm36_IND.gpkg")
```

So I think the best strategy for using this data is to start at the lowest level. Then you can also search across multiple geographies in a single table.

```{r}
india_gadm_3 <- st_read("./input_data/India/gadm36_IND.gpkg", layer="gadm36_IND_3") %>%
  janitor::clean_names()
```



### ind_88

This is a 2001 outbreak.

According to [this 2006 paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3373078/): “No definitive information about the possible index case exists.”

So, for now, let’s use the GADM layer to at least find the Siliguri area polygon.

Here we find what appears to be the correct polygon for Siliguri.

As you can see here, the select and mutate function calls are what create the end product mentioned in the intro section.

For precision, I've used "centroid" and "precise" --- with "precise" being the case when we are able to identify the exact location.

```{r}
ind_88 <- india_gadm_3 %>%
  filter(str_detect(name_3, "Siliguri")) %>%
  select(poly_geom = geom) %>% # this right here is where we start building the end dataset
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(id = "ind_88",
         the_year = 2001) %>%
  mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```


### ind_63

This is an outbreak that's been cited as occuring in Nadia, India. 

According to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3321761/): 

"During April 9–28, 2007, five persons became ill and died within a few days at village Belechuapara, Nadia district, West Bengal, India, which borders Bangladesh."

It appears the Walsh coordinates are exact for this record. 

```{r}
ind_63 <- tribble(
  ~latitude, ~longitude,
  23.812481, 88.570558
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>%
  select(point_geom = geometry) %>%
  mutate(precision = "precise",
         id = "ind_63",
         the_year = 2007) %>%
  mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```



## Bangladesh

Get GADM data for repeated use. 

```{r}
st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")
```

```{r}
bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg", layer="gadm36_BGD_4") %>%
  janitor::clean_names()

```


### bgd_87

[This paper identifies an outbreak](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010) in Meherpur, Bangladesh from April-May 2001. 

"The first NiV outbreak was reported in April 2001 from a village in district Meherpur, Bangladesh with 13 confirmed cases and 9 (69.2%) deaths."

In Bangladesh, the administrative breakdown goes:

* Central government (level 0)
* Divisions (level 1)
* Districts, aka "zila" (level 2)
* Sub-districts, aka "upazila" (level 3)
* Union councils, municipalities and city corporations (level 4)

The union councils are rural areas and villages fall beneath those. For now, we don't have a good villages dataset. 

The above paper leads us to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf), which says:

"Surveys took place in the villages of Chandpur (population 604), where persons who died or were hospitalized had resided..."

Walsh, M. G. (2015) place this location at 23.776652, 88.645198. Let's check if we can be more accurate. 

First let's create the point. 

```{r}
bgd_87_point <- tribble(
  ~latitude, ~longitude,
  23.776652, 88.645198
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```


Now, we can turn to [this official small-area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Meherpur.pdf) to find the smallest geography containing the village. See page 14. 

```{r}
bgd_87 <- bgd_gadm_4 %>%
  filter(name_2 == "Meherpur") %>%
  filter(name_4 == "Pirojpur")
```

Now let's compare the Walsh point to our current geography.

```{r}
ggplot() + 
  geom_sf(data=bgd_87) + 
  geom_sf(data=bgd_87_point, color="red")
```

We can see this is an improvement in geography. We'll use our new polygon. 

```{r}
bgd_87 <- bgd_gadm_4 %>%
  filter(name_2 == "Meherpur") %>%
  filter(name_4 == "Pirojpur") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(the_year = 2001, 
         id = "bgd_87") %>%
  mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```



### bgd_76

This outbreak in Naogoan is also profiled in [the same paper used for bgd_87]((https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf)), the outbreak in Chandpur village, Meherpur district. 

Unfortunately, the paper never tells us where the actual index case occured. Here's two relevant passages: 

"Surveys were conducted in the adjacent villages of East Chalksita (population 529) and Biljoania (population 481); suspected deaths and hospitalizations caused by Nipah virus infection were reported from both villages."

and 

"The index case occurred in a 12-year-old boy with symptom onset on January 11, 2003, and the last case occurred in 12-year-old girl on January 28 (Figure B)."

But they never actually tell us where the index case lived.  

In this case, we're once again left with a large area and unable to narrow it down. This needs additional reporting. 

In [this report from ICDDR,B](http://dspace.icddrb.org/jspui/bitstream/123456789/4811/1/2003-ICDDRBHealthScienceBulletin-Vol1%285%29-English.pdf), we learn:

"Between 11 and 28 January 2003, another outbreak of severe illness including
features of encephalitis was reported affecting at least 17 residents (range 4-
42 years) of Chalksita and Biljoania villages (located 45 km north-east of
Rajshahi) in Naogaon district; 8 people died."


Walsh used 24.814569, 88.927434, but that appears to be a centroid. 


```{r}
bgd_76_point <- tribble(
  ~latitude, ~longitude,
  24.807898, 88.638664
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```

On page 22 of [this small-area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Naogaon.pdf), we can see in the index that Chak Sita and Bil Joania are in Sreemantapur Union (in NIAMATPUR UPAZILA).

We can use that information to find the level 4 polygon in the GADM data.

FYI: The GADM dataset spells it Naogaon. 

```{r}
bgd_76 <- bgd_gadm_4 %>%
  filter(name_2 == "Naogaon") %>%
  filter(name_4 == "Sreemantapur")
```

Final record.

```{r}
bgd_76 <- bgd_gadm_4 %>%
  filter(name_2 == "Naogaon") %>%
  filter(name_4 == "Sreemantapur") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(the_year = 2003,
         id = "bgd_76") %>%
  mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```



http://dspace.icddrb.org/jspui/bitstream/123456789/4816/1/2004-ICDDRBHealthScienceBulletin-Vol2%282%29-English.pdf


### bgd_73

The MetaBiota data identifies a 2004 outbreak in Fardipur.

In this [WHO report](https://webcache.googleusercontent.com/search?q=cache:JHn6ZBsaUBQJ:https://apps.who.int/iris/rest/bitstreams/949669/retrieve+&cd=1&hl=en&ct=clnk&gl=uk), we learn the following:


Cluster of Nipah cases in Faridpur district, March–April 2004

"On 19 April, the Bangladesh Ministry of Health reported one cluster of 30 Nipah cases, including 18 deaths from 13 March to 14 April in Faridpur district. Laboratory testing, performed by CDC, confirmed Nipah virus infection in 16 of the cases.
The outbreak was investigated by a team comprising experts from the IEDCR, ICDDRB and the WHO country office. All cases lived in Guha Laksmipur village or were in contact with a spiritual leader living in this village who was himself a case. Direct contact with ill patients is suspected to have played a role in the transmission of the disease and the spread of this outbreak. The team organized infection control and safe clinical management workshops for local medical staff in different health care settings of the Faridpur district. Investigations are continuing."

On page 34 of this [Bangladeshi government atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Faridpur.pdf), we can identify these villages in the admin level 4 area of Sadar Paurashava.

This [clip seems to confirm](https://www.asianews.it/news-en/Faridpur:-Hindu-family-killed-by-bat-virus-17413.html) our placement

"In 2004, a NiV outbreak killed 17 people in the village of Laxmipur, Sadar upazila."

```{r}
bgd_73 <- bgd_gadm_4 %>%
  filter(name_2 == "Faridpur") %>%
  filter(name_4 == "Faridpur Paurashava") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(the_year = 2004,
         id = "bgd_73") %>%
  mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```


### bgd_72

This is an outbreak in Rajbari District, Bangladesh during January and February 2004. It occurred at the same time as another outbreak in Manikganj District, Bangladesh, which isn't in the Metabiota data, now discussed in the adds file at bgd_96. 

According to this paper:

"On 17 January 2004, a 9-year-old child from Kazipara village was brought to the emergency room of Goalanda health complex in Rajbari district, Bangladesh. The patient had a three-day history of fever and worsening neurological signs: he died within 15 minutes of admission. Between 17 and 21 January, 6 additional children (aged between 2 and 15 years) and a 28 year-old woman (mother of the youngest child) also died with the same symptoms. All patients came from the neighbouring Kazipara and Juran Molla Para villages and lived in two groups of houses located within 1 km of each other."

And we learn from [this 2008 paper](https://academic.oup.com/cid/article/46/7/977/292786) that it's in Goalanda subdistrict. 

So first, let's use GADM to try to identify the polygon for Goalanda subdistrict. GADM appears to spell Goalanda as Goalandaghat.


```{r}
bgd_72 <- bgd_gadm_4 %>%
  filter(name_3 == "Goalandaghat") %>%
  select(name_3) %>%
  group_by(name_3) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom) %>%
  mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```

We should try to nail down the location of Kazipara village.

By using the [Rajbari small area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Rajbari.pdf), we're able to figure out that Kazi Para, as its spelled in the document, is in GOALANDAGHAT PAURASHAVA. 

If we go back to our map, we can see that is a level 4 geography. While we don't know the exact village location, we've now narrowed it down to a tiny area. The Goalandaghat Paurashava is in red. 

So we're now going to use that as our polygon and take the centroid of that. 

```{r}
bgd_72 <- bgd_gadm_4 %>%
  filter(name_3 == "Goalandaghat") %>%
  filter(name_4 == "Goalandaghat Paurashava") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
    mutate(the_year = 2004,
         id = "bgd_72") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")

  
```

Absent the exact location of the village, this is very accurate. 



### bgd_68 

MetaBiota identified this 2005 outbreak in Tangail, Bangladesh.

The Deka data describes this outbreak as occuring at Bashail Upazilla, Tangail District, Bangladesh. 

[This bulletin indicates](http://dspace.icddrb.org/jspui/bitstream/123456789/4852/1/2005-ICDDRBHealthScienceBulletin-Vol3%284%29English.pdf) a location slightly more specific: "The study team defined a case of outbreak associated encephalitis as a person who lived or travelled in Habla Union, Basail Upazila, Tangail district, Bangladesh and developed fever with the new onset of seizures or altered mental status between 15 December 2004 and 31 January 2005."

Let's use GADM to try to identify the lowest-level polygon that matches our description.

Absent finding the exact village location, this is the most accurate spot we're going to get. 

```{r}
bgd_68 <- bgd_gadm_4 %>%
  filter(name_1 == "Dhaka") %>%
  filter(name_2 == "Tangail") %>%
  filter(name_4 == "Habla") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(the_year = 2005,
         id = "bgd_68") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```



### bgd_62

According to the original Metabiota data, this is a 2007 outbreak in Naoganon, site of a previous outbreak in 2003. However, I can find little information about this outbreak. It is also mentioned [in this USDA document](https://www.star-idaz.net/app/uploads/2018/09/Henipavirus-Gap-Analysis-Report-August-2018.pdf) (Pabna, Natore and Naogaon). A review of IECDR outbreak investigations done during 2007 reveals [no outbreak investigations for Naogaon during 2007](https://iedcr.gov.bd/outbreak/c4fb4e31-5bf4-4235-b4cf-f2dc5a710dda).

However, while investigating this event, I discovered we appear to be missing a 2007 Bangladesh outbreak in Sadar Upazilla, Kushtia District, Bangladesh. I have added it to the potential adds tab in the database as record x8.

For now, I'm going to leave this alone and move onto the next record. I will ask Metabiota if they have any other cites. 


### bgd_61

Metabiota lists this 2007 event as occuring in Thakurgaon.

IEDCR's list of 2007 outbreaks [lists](https://webcache.googleusercontent.com/search?q=cache:jJqzeuQUN48J:https://old.iedcr.gov.bd/website/index.php/outbreak/85-outbreak-list-of-2007+&cd=1&hl=en&ct=clnk&gl=ca) a Nipah outbreak in February in Thakurgaon, with no other details. 

According to the Deka spreadsheet, this outbreak occured in Haripur Upazilla, Thakurgaon District, Bangladesh. 

[This paper](https://www.cambridge.org/core/journals/epidemiology-and-infection/article/nipah-virus-outbreak-with-persontoperson-transmission-in-a-district-of-bangladesh-2007/7C468D7713F68FD41E7706B9AB0AA034) looks to describe the outbreak:
"On 9 February 2007, a physician at Rangpur Medical College Hospital, one of 10 hospitals involved in active NiV encephalitis surveillance in Bangladesh, reported a cluster of fatal encephalitis involving a husband and a wife residing in the Haripur Upazila (subdistrict) of Thakurgaon District."

Later there is a reference to "the outbreak village." And it says the index patient and some others "consumed raw date palm juice bought ... from a vendor in the local village market."

The researchers also sampled bat roots 1km and 15km from teh village: "The NiV neutralizing antibody prevalence was >50% in the bats sampled from the outbreak area which suggests that NiV has circulated in this population of bats."


```{r}
bgd_61 <- bgd_gadm_4 %>%
  filter(name_2 == "Thakurgaon") %>%
  filter(name_3 == "Haripur") %>%
  select(name_3) %>%
  group_by(name_3) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(the_year = 2007,
         id = "bgd_61") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```



### bgd_56

According to [this paper](https://www.ncbi.nlm.nih.gov/sites/books/NBK114490/):

"The communities affected by the outbreak were located in ... Baliakandi Upazila (population 186,562) of Rajbari District in central-western region of Bangladesh (Bangladesh Bureau of Statistics, 2010). These sites are located 44km apart and are separated by a river (Figure A12-1). Date palm sap is widely harvested, sold, and consumed in both areas."

Another key detail: 

"In the Rajbari cluster, three members from one family (mother and her two children) shared date palm sap purchased from the neighborhood date palm sap collector with two other neighborhood residents (brother-in-law and nephew of that mother) on February 18; all five subsequently developed illness. A salesman who resided nearly 5 km away from those households visited the village that morning and also drank the sap offered to him. He also died with the similar symptoms to the other four cases in Rajbari."

From [news coverage](https://www.thedailystar.net/news-detail-25894?amp), we learn: 

"The dead are Aleya Begum, 30, wife of Kamal Mollah of Borohijla village in Rajbari, their sons Jewel, 14, and Robiul Islam, 9."

On page 10 of the Bangladesh [small-area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Rajbari.pdf), we see that a village called Bara Hijli is in Nawabpur Union. 

```{r}
bgd_56 <- bgd_gadm_4 %>%
  filter(name_2 == "Rajbari") %>%
  filter(name_3 == "Balia Kandi") %>%
  filter(name_4 == "Nawabpur") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
    mutate(the_year = 2008,
         id = "bgd_56") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
  
```

Absent a specific village coordinate, this appears to be the most accurate conclusion. 



### bgd_55

This outbreak occured at the same time as bgd_56 and the two areas are very close to each other. 

According to [this paper](https://www.ncbi.nlm.nih.gov/sites/books/NBK114490/):

"In the Manikgonj cluster, three children from one family drank raw date palm sap, collected by their father, a local gachi or date palm sap collector. They drank the sap early in the morning on February 11 for the last time and subsequently two of them developed illness on February 20 and the third on February 23. The fourth child, who developed illness on February 21, was a resident of Dhaka District but visited his grandmother’s house on February 6 for 12 days. His grandmother, a neighbor of the date palm sap collector, purchased raw date palm sap from him and served it to her grandson the same day the other children
drank the sap."

According to [this article](https://www.thedailystar.net/news-detail-26658):

"IEDCR confirmed the first spell of the Nipah outbreak on February 29 based on the samples collected from the bodies of the three children who hailed from Bishnupur village under Daulatpur Upazila in Manikganj."

We can [see on page 10](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Manikganj.pdf), Bishnupur village is in the Khalsi Union (level 4) geography.


```{r}
bgd_55 <- bgd_gadm_4 %>%
  filter(name_1 == "Dhaka") %>%
  filter(name_2 == "Manikganj") %>%
  filter(name_3 == "Daulatpur") %>%
  filter(name_4 == "Khalsi") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
    mutate(the_year = 2008,
         id = "bgd_55") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```



### bgd_50

This was a 2009 outbreak in Rangpur. 

Deka did not cite this in his data. Can we access Walsh supplementary material to see if they had it in their original paper?

For now, unfortunately, I'm going to leave this as-is. 

```{r}
bgd_50 <- bgd_gadm_4 %>%
  filter(name_1 == "Rangpur") %>%
  filter(name_2 == "Rangpur") %>%
  select(name_2) %>%
  group_by(name_2) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(the_year = 2009,
         id = "bgd_50") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
  
```





### bgd_49

This was a 2009 outbreak in Rajbari. 

I don't know if it's Rajbari district or town or what??? Can't find anything anywhere. 

Come back to this. 

### bgd_45

Metabiota lists this as a 2010 outbreak in Dhaka. Walsh has no info on this outbreak. 

According to [this bulletin](http://dspace.icddrb.org/jspui/bitstream/123456789/4870/1/2010-ICDDRBHealthScienceBulletin-Vol8%282%29-English.pdf):

"The index case was a 45 year old male from Bhanga sub-district who died after being admitted to the hospital. Another 40 year old female case, who was a neighbour of the index case, presented with vomiting, headache, convulsion, altered mental status, and loss of consciousness, and died at home. Both of them were probable cases. Two other cases from the 40-year old female’s family,
10 and 11 year old girls, who were admitted in hospital with a clinical diagnosis of encephalitis, died. One girl was found anti-Nipah IgM positive. The other girl was a probable Nipah case."

To be clear, the paper points out this is in Dhaka division, Faridpur district. 

Finally, [ProMed tells us](https://promedmail.org/promed-posts/):

"The deceased were a 30 year old woman of Algi village under Bhanga upazila, her 11 year old daughter, and 10 year old nephew. Family sources said the woman was the wife of a fisherman and she died on 13 Jan 2010 after she had a fever and headache. The same day her daughter was admitted to Bhanga Health Complex with the same symptoms. Later she was transferred to Faridpur Medical College Hospital where she died on 16 Jan 2010. Her nephew also died of the same disease on 17 Jan 2010 at the Faridpur Medical
College Hospital. All the 3 had consumed raw date juice before they contracted the disease, family sources added."

IEDCR's list of outbreak investigations for that year also lists the village as Algi, [here](https://old.iedcr.gov.bd/website/index.php/outbreak/88-list-of-outbreak-investigations-in-2010-updated-31-december-2010)


```{r}
bgd_45 <- bgd_gadm_4 %>%
  filter(name_1 == "Dhaka") %>%
  filter(name_2 == "Faridpur") %>%
  filter(name_3 == "Bhanga") %>%
  filter(name_4 == "Algi")
```

Let's compare to the Walsh dot. 

```{r}
bgd_45_point <- tribble(
  ~latitude, ~longitude,
  23.381598, 90.016138
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```

From the looks of this, it appears we have improved on the location.

```{r}
ggplot() + 
  geom_sf(data=bgd_45) + 
  geom_sf(data=bgd_45_point)
```


```{r}
bgd_45 <- tribble(
  ~latitude, ~longitude,
  23.381598, 90.016138
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) %>%
  select(point_geom = geometry) %>%
  mutate(the_year = 2010,
         id = "bgd_45") %>%
    mutate(family = "Paramyxoviridae",
         genus = "Henipavirus",
         species = "Nipah henipavirus")
```




### bgd_44

Metabiota identified this 2011 outbreak in Comilla, Dinajpur, Faridpur, Lalmohirhat, Nilphamari. 

Deka's data identifies an outbreak at Hatibandha, Lalmonirhat, Bangladesh. This is confirmed via a [Nipah update published by IEDCR](http://www.iedcr.org/pdf/files/nipah/Nipah-Update.pdf). 

We may actually have multiple outbreaks, [according to this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4675679/).

We will focus on Lalmonirhat outbreak here and code other districts in the add table. 

From that paper, covering a period from December 23, 2010 to March 1, 2011: "There were 37 cases in four clusters: three clusters occurred in three northern districts, Lalmonirhat (22 cases), Rangpur (eight cases), and Dinajpur (five cases), and the fourth cluster was in Rajbari District (two cases), in central Bangladesh (Fig. 1). All four clusters were detected through hospital-based Nipah surveillance. The remaining six cases were isolated, five of which were detected through hospital-based surveillance."

It describes the Lalmonirhat outbreak as one cluster: 
"Of the 22 cases in the Lalmonirhat cluster, 21 lived in five adjacent villages and the other case was a healthcare worker in a referral hospital where cases from this cluster were admitted for treatment."

[This news clip](https://www.thedailystar.net/news-detail-172896) describes the situation thusly: 

"The dead were identified as Bachchu Miah, 56, of Baraipara village; Joynal, 42, of South Goddimari; and Asadul Islam, 18, of Gendukuri village in Hatibandha upazila of Lalmonirhat and Lubna, 14, of South Kolkan village in Gangachara upazila of Rangpur.
Of them, Bachchu and Joynal died at Hatibandha Upazila Health Complex at about 11:30pm on Thursday and Asadul died at Rangpur Medical College Hospital early hours yesterday and Lubna at her residence around 12:00noon."

However, based on this description, it's tough to determine the index case and the best we can do for now is level 3.

Some good color on public health measures [here](https://arrkhan.blogspot.com/2011/02/schools-closed-to-contain-nipah.html): "Authorities in Lalmonirhat have closed down schools, madrasahs and other educational institutions in a bid to prevent the spread of the disease."

More color from a New Nation piece from Feb 5, 2011 archived by Factiva: "People are still deserting their homes in panic while the local authorities shut down all primary schools at Hatiabandha upazila, which has witnessed the recent outbreak.
"Even ambulance drivers fled the village. We are trying to convince people aking them not to panic and stop drinking raw palm or date sap," said Dr Farhana Haque, an investigator of the government's team despatched to the upazila on February 2 soon after the outbreak.
[...]
Anthropologists say people generally believe that evil spirit has entered their community causing the illness.
They interpreted the illness as 'asmani bala' (divine burden). Some people of the community even have the perception that bats cannot contaminate raw date or palm sap. And do not believe it could be transmitted from person to person.
"We have to develop a rapport and earn their trust. We let them to talk first and give them space to express their emotions or complaints," said ICDDR,B anthropologist Shahana Parveen.
"With the respect of cultural norms and local practice, we make people understand what it actually is and what they should do."
"We held a courtyard meeting today to sensitise the idea that germs were responsible for the illness and deaths, using local terms and familiar examples," Farhana Haque said, adding that the participants responded positively after being given an explanation with evidence."


```{r}
bgd_44 <- bgd_gadm_4 %>%
  filter(name_1 == "Rangpur") %>%
  filter(name_2 == "Lalmonirhat") %>%
  filter(name_3 == "Hatibandha") %>%
  select(name_3) %>%
  group_by(name_3) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid")
  

```




"Of the eight cases in the Rangpur cluster, proxies of three probable cases reported that the cases regularly drank fermented date palm sap, a traditional liquor known locally as tari, from a common source. They consumed the drink daily during the 30 days prior to their illness onset. These probable cases did not have any history of drinking raw sap or of having contact with a Nipah case. Three confirmed cases became ill following close contact with one of these probable cases. Two other cases from this cluster, one confirmed and one probable, had no reported history of consuming raw or fermented date palm sap or contact with a Nipah case."



# Assembly 

```{r}
ryan_nipah_unfinished <- bind_rows(ind_88, ind_63, bgd_87, bgd_76, bgd_73, bgd_72, bgd_68, bgd_61, bgd_56, bgd_55, bgd_50, bgd_45)  
```

Clean up after ourselves.

```{r}
rm(ind_88, ind_63, bgd_87, bgd_76, bgd_73, bgd_72, bgd_68, bgd_61, bgd_56, bgd_55, bgd_50, bgd_45)
```

```{r}
rm(bgd_45_point, bgd_76_point, bgd_87_point, bgd_gadm_4, india_gadm_3)
```

Save it to disk.

```{r}
write_rds(ryan_nipah_unfinished, "./spatial_outputs/ryan_nipah_unfinished.rds")
```





```{r}
st_geometry(ryan_nipah_unfinished) <- "point_geom"
```


```{r}
st_write(ryan_nipah_unfinished, "./spatial_outputs/ryan_nipah_unfinished.gpkg", append=FALSE)
```