---
title: "Nipah sites in Bangladesh"
output:
  github_document:
    toc: true
---

```{r}

library(tidyverse)
library(sf)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)

```

Set working directory for Allison's Windows laptop

```{r}

setwd("C:/Users/0146156/Documents/mb_checks/reuters-news-data_disease_gis")

```

Get GADM data. 

```{r}

st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")

bgd_gadm_2 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_2") %>%
              janitor::clean_names()

bgd_gadm_3 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_3") %>%
              janitor::clean_names()

bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer="gadm36_BGD_4") %>%
              janitor::clean_names()

st_layers("./input_data/India/gadm36_IND.gpkg")

ind_gadm_3 <- st_read("./input_data/India/gadm36_IND.gpkg",
                      layer = "gadm36_IND_3") %>%
              janitor::clean_names()

```

Get other spatial data.

```{r}

st_layers("./input_data/India/Kerala")

ind_taluks <- st_read("./input_data/India/Kerala", layer = "taluk") %>%
  rename(geom = geometry, oid = OID_, taluk = TALUK, state = STATE, district = DISTRICT)

st_layers("C:/Users/0146156/Documents/mb_checks/stanford_level_3")

phl_level_3 <- st_read("C:/Users/0146156/Documents/mb_checks/stanford_level_3")

```

Setting up function that returns variables based on single GADM polygon.

```{r}

grab_polygon <-function(geom, id) {
  geom %>% select(poly_geom = geom) %>% 
           mutate(point_geom = st_centroid(poly_geom),
                  precision = "centroid",
                  reuters_id = id,
                  country_iso = "BGD")
}

```

# INDIA 

#ind_7

Metabiota lists Kerala, 2019, no other details, but Deb has market coordinates from original reporting.

```{r}

temp_point <- tribble(~latitude, ~longitude, 10.17222, 76.21215) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

ind_7 <- data.frame(poly_geom = NA,
                    point_geom = temp_point,
                    precision = "point",
                    reuters_id = 7,
                    country_iso = "IND")

```
#ind_11

Kozhikode, 2018, no other details. Kozhikode is a district and a city in Kerala.

There is an index case for this outbreak: "An exhaustive study by the Kerala Health Department tracing the ‘contact history’ of patients affected by the recent Nipah outbreak in the state has found that 17 of the 19 infected persons got the virus directly from Patient No. 1, Muhammed Sabith. [...] Sabith is believed to have caught the infection sometime in the third week of April. On May 2, he developed fever and severe body pain. He went to a nearby hospital in Perambra to get examined, after which he was sent home."
[source](https://indianexpress.com/article/india/tracking-nipah-outbreak-sabith-patient-no-1-infected-all-but-one-5259668/)

Long feature confirms that Sabith lived in Perambra town, and first went to Perambra Taluk Hospital. [source](https://www.onmanorama.com/news/kerala/2018/10/02/nipah-virus-causing-bats-ate-fruits-probe-kozhikode.html)

Perambra Taluk Hospital is at (11.569284, 75.755022). Wikipedia says Perambra is in the Koyilandy taluk of Kozhikode district, and Koyilandy can also be transliterated as Quilandy. Coding this case to Quilandy.

```{r}

# code to Quilandy taluk polygon
ind_11 <- ind_taluks %>% filter(str_detect(taluk, "Quilandy")) %>%
  grab_polygon(11)

# load world map data
world <- ne_countries(scale = "medium", returnclass = "sf")

# create shapefile with point for hospital where index patient went first
hospital <- tribble(~latitude, ~longitude, 11.569284, 75.755022) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326)

# map taluk and hospital location to make sure hospital is within the taluk
ggplot(world) +
  geom_sf() +
  geom_sf(data = ind_11$poly_geom, aes(color = "blue")) +
  geom_sf(data = hospital, aes(color = "red")) + 
  coord_sf(xlim = c(74, 80),
           ylim = c(11, 13))

# Calculate distance between the hospital and the centroid of this taluk
st_distance(hospital, ind_11$point_geom)


# After discussion with Ryan, replacing centroid point geometry with hospital location and creating a new precision category
ind_11 <- ind_11 %>% mutate(point_geom = hospital, precision = "hospital") %>% str

```
# PHILIPPINES

#phl_22

2014, area listed as "Sultan Kudarat Senator Ninoy Aquino", no other details.

Academic paper is [here](https://wwwnc.cdc.gov/eid/article/21/2/14-1433_article)

"On April 2, 2014, the Philippine National Epidemiology Center received a report of human deaths in 2 villages, Tinalon and Midtungok, in the municipality of Senator Ninoy Aquino, province of Sultan Kudarat, island of Mindanao. The villages are ≈15 km apart, and the provincial referral hospital is in Isulan, 80 km away."

Likely spillover: "Although the source of the horse infections is unclear, on the basis of the known ecology of henipaviruses, the most likely source is fruit bats (family Pteropodidae) (10). Bats belonging to this family were reported near at least 1 of the 2 villages."

It is not completely established that this was Nipah and not Hendra but there seem to be good reasons to classify it as Nipah: "Clinical presentations, epidemiologic findings, and serologic results suggest that the virus causing this outbreak was a henipavirus. It was most likely NiV or a virus that is antigenically and genetically closely related to NiV."

```{r}

phl_22 <- phl_level_3 %>% filter(str_detect(name_3, "Tinalon")) %>%
  rename(geom = geometry) %>%
  grab_polygon(22) %>% 
  mutate(country_iso = "PHL")

```


# BANGLADESH

# bgd_3

Unspecified number of Nipah cases in Bangladesh, 2021, no other details in data.

ProMedMail points to a short Outbreak News piece about an updated Nipah case count for Bangladesh in 2021, listing two cases [here](http://outbreaknewstoday.com/bangladesh-reports-two-nipah-virus-cases-in-2021/)

That is based on this chart from Bangladesh's Institute of Epidemiology, Disease Control and Research, [here](https://iedcr.gov.bd/surveillances/93c87e70-9c22-4f21-9506-a1161ecf404f) 

No way to pinpoint potential spillover location based on this source. 

@waiting AM emailed IEDCR director about location of cases from their statistics, Jan 11
*** For now, leaving this data point out 

### bgd_5

Unspecified number of Nipah cases in Bangladesh in 2020, no other details.

Likely the same source as above but unclear. 

Searched ProMed

@waiting IEDCR director
*** For now, leaving this data point out 

### bgd_6

Unspecified number of Nipah cases in Thakurgaon District in 2019.

Academic paper describes this outbreak, with cases from Jan 15 to March 17 [here](https://www.ijidonline.com/action/showPdf?pii=S1201-9712%2820%2932414-0)

Primary case: "He was a traditional healer, had no history of taking RDPS, occasionally visited forest to collect herbs and animals."

No additional information about location here. 

Mention in this local press report of Baliadangi Upazila Health Complex, in relation to March cases, but paper also notes "In February, five members of a family died mysteriously in Baliadanga upazila" [here](https://www.newagebd.net/article/67534/3-of-a-family-suspected-to-be-infected-with-nipah-virus)

Seems likely that IEDCR can also sort this out. 

*** Coding to Baliadangi Upazila

```{r}

bgd_6 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Baliadangi")) %>%
  grab_polygon(6)

```

#bgd_14 to bgd_19

These are all clearly based on an Outbreak News summary of cases in 2015 reported by IEDCR by district [here](http://outbreaknewstoday.com/bangladesh-reports-nine-nipah-virus-cases-to-date-in-2015/)

Nothing further on ProMedMail.

Local media paces a Naogaon cluster (bgd_17) in "Valain village under Manda Upazila" [here](https://www.risingbd.com/english/Nipah_virus_kills_3_in_Naogaon/22046)

Daily Star (article)[https://www.thedailystar.net/2-die-of-nipah-virus-in-naogaon-62907]

May 18, 2015: "Two people died of Nipah virus at Manda upazila in Naogaon last week, said the Institute of Epidemiology, Disease Control and Research today."

@waiting IEDCR director
*** Coding bgd_17 to Manda Upazila, the rest at the district level

```{r}

bgd_14 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Faridpur")) %>%
  grab_polygon(14)

bgd_15 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Rajbari")) %>%
 bgd_17 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Manda")) grab_polygon(15)

bgd_16 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Magura")) %>%
  grab_polygon(16)

bgd_18 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Nilphamari")) %>%
  grab_polygon(18)

bgd_19 <- bgd_gadm_2 %>%
  filter(str_detect(name_2, "Panchagarh")) %>%
  grab_polygon(19)


```

# BANGLADESH, 2013

Full IECDR location list for 2013:
[source](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11)
Rajbari - district
Jhenaidah - district
Naogaon - district
Nator - probably Natore, district
Natore - district
Gaibandha - distict
Mymensingh - city and division
Pabna - city, district
Rajshahi - division, district and city
Nilphamari - district
Kurigram - district
Kushtia - district
Magura - district
Manikganj - district


ProMED-mail post about 2012/2013 season: 

"Among the casualties, there have been 2 from each of Dhaka, Pabna, and Natore and one from each of Rajbari, Jhenaidah, Naogaon, and Rajshahi, according to Bangladesh's Institute of Epidemiology Diseases Control and Research (IEDCR)."

Pabna (city and district)
Natore (district)
Rajbari (district)
Jhenadidah (district)
Naogaon (district)
Rajshahi (city, district and division)

[link](https://promedmail.org/promed-post/?id=1530748)

#bgd_24

Data says Dhaka, 2013, no other details.

Local press article from early 2013 notes first ever Nipah death in Dhaka, which is outside the Nipah belt, with this detail: "'Although the sap was brought from Mymensingh [District], it was consumed in Dhaka. This is the first case of a Nipah-related death in the city,' said Alamgir."
[source](https://reliefweb.int/report/bangladesh/fresh-outbreak-nipah-virus-bangladesh)

Earlier report from January: "According to its Director, the family drank raw date sap brought from Bhaluka on Jan 11 and fell ill six days later" - goes on to not ethat one person in Bhaluka supplied 100 bottles to families in Dhaka.

Bhaluka is an upazila within the Mymensingh district, but Bhaluka is also a village in the upazila. 

*** Coding to the upazila

```{r}

bgd_24 <- bgd_gadm_3 %>%
  filter(str_detect(name_3, "Bhaluka")) %>%
  grab_polygon(19)

```

#bgd_25

Data notes district Faridpur and 2013, no other details.

Nothing on ProMED or online that I can find.

Faridpur is not mentioned in EICDR's last 2013 report, though it was from May 2013 so possibly [incomplete](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11)

*** Not coding yet

#bgd_26

Data notes district Madaripur and 2013, no other details.

Nothing on ProMED or online that I can find.

Faridpur is not mentioned in EICDR's last 2013 report, though it was from May 2013 so possibly [incomplete](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11)

*** Not coding

#bgd_27

Data sais "Makikgank" which is not a district, but district Manikganj did have a documented outbreak that year:

"At Hafeznagar village in Shibalaya upazila of the district, 4 people, including a couple, died of an unknown disease in the last 2 weeks. The victims are a 40-year-old woman, her husband age 45 years, the woman's mother of 55 years, and a neighbour, a man 36-years-old. The neighbor and the woman died at Dhaka Community Hospital on 28 and 29 Mar [2013]. The mother died on way to Dhaka on 28 Mar [2013], while the husband died at Monno General Hospital in Manikganj on 14 Mar [2013], family sources said." [source](https://promedmail.org/promed-post/?id=20130401.1613261)

Follow up post confirms this was Nipah, see [here](https://promedmail.org/promed-post/?id=1615535)

Hafeznagara is not in Google Maps or the [Manikganj small area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Manikganj.pdf)

*** Coding to Shibalaya upazila

```{r}

bgd_27 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Shibalaya")) %>%
  grab_polygon(27)

```

#bgd_28

Data notes district Shariatpur and 2013, no other details.

Nothing in ProMED, Daily Star, general Google search, Factiva, IECDR report from May 2013.

*** Not coding 

#bgd_29 

Data notes district Kushtia and 2013, no other details.

*** Coding to Kushtia district

```{r}

bgd_29 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Kushtia")) %>%
  grab_polygon(29)

```
#bgd_30

Magura district, 2013, no other details.

One case is described in ProMED:
"Nipah virus claimed the life of a minor boy at Ghosgati village in Shalikha upazila on Sunday [10 Mar 2013]. A boy, aged 2, of the village died at Jessore General Hospital in the morning. District Civil Surgeon Dr Muhammad Shafiqul Alam said the boy had fallen sick after drinking date [palm] juice on Thursday [7 Mar 2013]. He was admitted to Magura district hospital."
[source](https://promedmail.org/promed-post/?id=1592574)

Magura only appears once on the IECDR list for the year [here](http://www.iedcr.org/index.php?option=com_content&view=article&id=135:23-rd-february-2013-nipah-outbreak&catid=11) so this is likely the right case

The small area atlas lists a "Ghoshgati" in Gangarampur Union of Shalikha upazila. Mindat lists coordinates for a Ghoshgati in Khulna but that's in the wrong upazila. 

*** Coding to Gangarampur Union

```{r}

bgd_30 <- bgd_gadm_4 %>% filter(str_detect(name_2, "Magura")) %>%
  filter(str_detect(name_4, "Gangarampur")) %>%
  grab_polygon(30)

```
#bgd_31

Chapainawabganj district, 2013 no other details.

Alternate transliteration is Chapai Nawabganj.

The district is not on IECDR's 2013 case list but it is on the 2014 one, which actually covers December 2013 as well: [here](http://www.iedcr.org/pdf/files/nipah/Nipah%20situation_18.02.2014.pdf)

Nothing in the Daily Star, Factiva, ProMED.

*** Not coding yet

#bgd_32

Naogaon district, 2013, no other details.

Naogaon is on IECDR's list of 2013 cases so at minimum we can code to the district.

Nothing further in Factiva or ProMED for that year. 

Daily Star mentions a death in Naogaon in January in passing [here](https://www.thedailystar.net/news-detail-266732): "The dead are from Dhaka, Rajbari, Jhenidah, Naogaon, Natore and Gaibandha."

*** Coding to Naogaon district.

```{r}

bgd_32 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Naogaon")) %>%
  grab_polygon(32)

```

#bgd_33

Natore district, 2013, no other details.

Natore is on IECDR's list of 2013 cases so at minimum we can code to the district. Nothing further in Factiva or ProMED for that year. 

Daily Star, Jan 27, 2013: "In Natore, a mother passed away while her eight-month-old baby has been infected." [source](https://www.thedailystar.net/news-detail-266732)

Similar from bdnews24.com [here](https://bdnews24.com/business/2013/01/25/nipah-claims-5-more-lives)

*** Coding to Natore district

```{r}

bgd_33 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Natore")) %>%
  grab_polygon(33)

```

#bgd_34

Listed as "Rjshahi", likely typo for Rajshahi, 2013, no other details. Rajshahi is a division, district and city. 

Rajshahi is on IECDR's list of 2013 cases. 

bdnews24.com, Jan 27, 2013: "The Director of the government’s disease monitoring outfit, IEDCR, Prof Mahmudur Rahman on Sunday said two more died of the virus taking the total death toll this year to eight. The new deaths were from Pabna and Rajshahi while one 8-month old boy is struggling for life in the Rajshahi Medical College Hospital."
[source](https://bdnews24.com/health/2013/01/27/nipah-virus-claims-2-more-lives)

Later story makes clear that the 8 month old is the one from Natore, see [here](https://bdnews24.com/bangladesh/2013/01/27/boy-detected-with-nipah-virus)

*** Coding to Rajshahi district

```{r}

bgd_34 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Rajshahi")) %>%
  grab_polygon(34)

```

#bgd_35

Dinajpur, 2013, no other details. Dinajpur is both a city and a district. It is not on IECDR's 2013 case list. Nothing further on ProMED, Factiva, DAily Star or BDNews.

*** Not coding

# bgd_36

Ranpur, 2013, no other details. Rangpur is a city, district and division, and it is not on EICDR's 2013 outbreak list. Nothing furhter on ProMED, Factiva, DAily 

*** Not coding

# bgd_41

Joypurhat, 2012, Metabiota data links to [this paper](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010)

The paper contains a table that lists 12 cases and 10 deaths in several districts including in January 2012.

Daily Star, January 26, 2012: "Panic has gripped the residents of different villages in Khetlal and Sadar upazilas of Joypurhat district following death of four children infected with “Nipah virus” in three days from January 20 to January 22. 
The dead are Anik Mondol, 7, son of Alam Mondol, Marium Begum, 7, daughter of Jahangir Hossain of Natunhaat Sardarpara in Joypurhat town, and Tohidul Islam, 8, son of Anwar Hossain of Dadra Jointigram, and Shakila Begum, 9, daughter of of Ikargara village in Khetlal upazila.
Anik died at Bogra Medical College and Hospital on January 20, Shakila at Joypur Sadar Hospital on January 21 and Marium and Tohidul at Rajshahi Medical College and Hospital on January 22."
[source](https://www.thedailystar.net/news-detail-219894)

Places mentioned above:
- Joypurhat town
- Ikagara village, Khetlal upazila

FluTrackers thread quotes stories no longer available online, and apparently on January 22 they said of Touhid and Marium both "fell sick after drinking date juice from a same tree." The thread is [here](https://flutrackers.com/forum/forum/emerging-diseases-other-health-threats-alphabetical-i-thru-z/nipah-virus/121571-bangladesh-nipah-virus-kills-30-in-joypurhat).

FluTrackers  also quotes an IECDR update that is no longer available, posted to teh thread Jan 23: "From 20 January 2012, two clusters of Nipah were identified in the Joypurhat Municipality area and Khetlal upazila of Joypurhat district. A joint team of IEDCR and ICDDR,B are working in the field now. Till date (23 rd January 2012), the team identified 3 cases from the municipal area and 2 cases from Khetlal. All 5 cases have died. Laboratory tests for Nipah were done in IEDCR. All four samples collected from suspected cases were positive for Nipah."

BDNews24 quotes from the same IEDR bulletin: "The virus that killed five people at Jaipurhat in the last four days has been identified as Nipah. The Institute of Epidemiology Diseases Control and Research (IEDCR) confirmed on its website on Tuesday that the five had indeed died from Nipah. A team from the institute are now working in the affected area. The IECDR said human infections took place in Joypurhat municiplatiy and Khetlal Upazila." [link](https://bdnews24.com/bangladesh/2012/01/24/nipah-virus-attack-confirmed-5-die)

Separately, BD News24: "This year, while IEDCR is grappling to check the spread in Joypurhat district's Sadarpara and Khetlal upazillas, where five persons died in this month's outbreak, the district civil surgeon says new suspects of infection are emerging from the area." 

*** Coding to Joypurhat Paurashava union (paurashava is Bengali for 'municipality')
*** Added second location, Khetlal Upazila, to 'possible adds' sheet

```{r}

bgd_41 <- bgd_gadm_4 %>% filter(str_detect(name_4, "Joypurhat")) %>%
  grab_polygon(41)

```

