---
title: "Nipah sites in Bangladesh"
output:
  github_document:
    toc: true
---

# TEST COMMIT JAN 10 

```{r}
library(tidyverse)
library(sf)
```

Get GADM data. 

```{r}
st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")
```

```{r}
bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg", layer="gadm36_BGD_4") %>%
  janitor::clean_names()

```


### bgd_87

[This paper identifies an outbreak](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010) in Meherpur, Bangladesh from April-May 2001. 

"The first NiV outbreak was reported in April 2001 from a village in district Meherpur, Bangladesh with 13 confirmed cases and 9 (69.2%) deaths."

What we don't know which Meherpur, though. There's Meherpur at levels 2, 3 and 4. 


```{r}
bgd_gadm_4 %>%
  filter(str_detect(name_3, "Naogaon"))
```
The above paper leads us to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf), which says:

"Surveys took place in the villages of Chandpur (population 604), where persons who died or were hospitalized had resided..."

So that's a clue. 