---
title: "Nipah sites in Bangladesh"
output:
  github_document:
    toc: true
---

```{r}

library(tidyverse)
library(sf)

```

Set working directory for Allison's Windows laptop

```{r}

#setwd("C:/Users/0146156/Documents/mb_checks/reuters-news-data_disease_gis")

```

Get GADM data. 

```{r}

st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")

```

```{r}
bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg", layer="gadm36_BGD_4") %>%
  janitor::clean_names()

```

### bgd_3

Unspecified number of Nipah cases in Bangladesh, 2021, no other details in data.

ProMedMail points to a short Outbreak News piece about an updated Nipah case count for Bangladesh in 2021, listing two cases: 
http://outbreaknewstoday.com/bangladesh-reports-two-nipah-virus-cases-in-2021/

That is based on this chart from Bangladesh's Institute of Epidemiology, Disease Control and Research: https://iedcr.gov.bd/surveillances/93c87e70-9c22-4f21-9506-a1161ecf404f 

No way to pinpoint potential spillover location based on this source. 

@waiting AM emailed IEDCR director about location of cases from their statistics, Jan 11

### bgd_5

Unspecified number of Nipah cases in Bangladesh in 2020, no other details.

Likely the same source as above but unclear. 

Searched ProMed




### bgd_87

[This paper identifies an outbreak](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010) in Meherpur, Bangladesh from April-May 2001. 

"The first NiV outbreak was reported in April 2001 from a village in district Meherpur, Bangladesh with 13 confirmed cases and 9 (69.2%) deaths."

What we don't know which Meherpur, though. There's Meherpur at levels 2, 3 and 4. 


```{r}
bgd_gadm_4 %>%
  filter(str_detect(name_3, "Naogaon"))

```
The above paper leads us to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf), which says:

"Surveys took place in the villages of Chandpur (population 604), where persons who died or were hospitalized had resided..."

So that's a clue. 