---
title: "Nipah sites notebook"
output:
  github_document:
    toc: true
---

```{r}
library(tidyverse)
library(sf)
```


## India

### ind_88

According to [this 2006 paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3373078/): “No definitive information about the possible index case exists.”

So, for now, let’s use the GADM layer to at least find the Siliguri area polygon.

First, let’s check the layers.

```{r}
st_layers("./input_data/India/gadm36_IND.gpkg")
```

So I think the best strategy for using this data is to start at the lowest level. Then you can also search across multiple geographies in a single table.

```{r}
india_gadm_3 <- st_read("./input_data/India/gadm36_IND.gpkg", layer="gadm36_IND_3") %>%
  janitor::clean_names()
```

Here we find what appears to be the correct polygon for Siliguri.

```{r}
ind_88 <- india_gadm_3 %>%
  filter(str_detect(name_3, "Siliguri")) %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid")
```

