---
title: "Nipah sites in Bangladesh"
output:
  github_document:
    toc: true
---

```{r}

library(tidyverse)
library(sf)

```

Set working directory for Allison's Windows laptop

```{r}

#setwd("C:/Users/0146156/Documents/mb_checks/reuters-news-data_disease_gis")

```

Get GADM data. 

```{r}

st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")

```

```{r}

bgd_gadm_2 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_2") %>%
              janitor::clean_names()

bgd_gadm_3 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_3") %>%
              janitor::clean_names()

bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer="gadm36_BGD_4") %>%
              janitor::clean_names()

```

### bgd_3

Unspecified number of Nipah cases in Bangladesh, 2021, no other details in data.

ProMedMail points to a short Outbreak News piece about an updated Nipah case count for Bangladesh in 2021, listing two cases [here](http://outbreaknewstoday.com/bangladesh-reports-two-nipah-virus-cases-in-2021/)

That is based on this chart from Bangladesh's Institute of Epidemiology, Disease Control and Research, [here](https://iedcr.gov.bd/surveillances/93c87e70-9c22-4f21-9506-a1161ecf404f) 

No way to pinpoint potential spillover location based on this source. 

@waiting AM emailed IEDCR director about location of cases from their statistics, Jan 11
*** For now, leaving this data point out 

### bgd_5

Unspecified number of Nipah cases in Bangladesh in 2020, no other details.

Likely the same source as above but unclear. 

Searched ProMed

@waiting IEDCR director
*** For now, leaving this data point out 

### bgd_6

Unspecified number of Nipah cases in Thakurgaon District in 2019.

Academic paper describes this outbreak, with cases from Jan 15 to March 17 [here](https://www.ijidonline.com/action/showPdf?pii=S1201-9712%2820%2932414-0)

Primary case: "He was a traditional healer, had no history of taking RDPS, occasionally visited forest to collect herbs and animals."

No additional information about location here. 

Mention in this local press report of Baliadangi Upazila Health Complex, in relation to March cases, but paper also notes "In February, five members of a family died mysteriously in Baliadanga upazila" [here](https://www.newagebd.net/article/67534/3-of-a-family-suspected-to-be-infected-with-nipah-virus)

Seems likely that IEDCR can also sort this out. 

@waiting IEDCR director
*** Coding to Baliadangi Upazila

```{r}

bgd_6 <- bgd_gadm_3 %>% 
  filter(str_detect(name_3, "Baliadangi")) %>%
  select(poly_geom = geom) %>% 
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(reuters_id = 6,
         country_iso = "BGD")

```

### bgd_14 to bgd_19

These are all clearly based on an Outbreak News summary of cases in 2015 reported by IEDCR by district [here](http://outbreaknewstoday.com/bangladesh-reports-nine-nipah-virus-cases-to-date-in-2015/)

Nothing further on ProMedMail.

Local media paces a Naogaon cluster (bgd_17) in "Valain village under Manda Upazila" [here](https://www.risingbd.com/english/Nipah_virus_kills_3_in_Naogaon/22046)

Next: search Nilphamari

```{r}



bgd_gadm_3 %>%
  filter(str_detect(name_2, "Nilphamari"))


bgd_gadm_4 %>%
  filter(str_detect(name_2, "Nilphamari"))


bgd_gadm_4 %>% str


```

### bgd_87

[This paper identifies an outbreak](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010) in Meherpur, Bangladesh from April-May 2001. 

"The first NiV outbreak was reported in April 2001 from a village in district Meherpur, Bangladesh with 13 confirmed cases and 9 (69.2%) deaths."

What we don't know which Meherpur, though. There's Meherpur at levels 2, 3 and 4. 


```{r}
bgd_gadm_4 %>%
  filter(str_detect(name_3, "Naogaon"))

```
The above paper leads us to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf), which says:

"Surveys took place in the villages of Chandpur (population 604), where persons who died or were hospitalized had resided..."

So that's a clue. 