---
title: "Nipah sites notebook"
output:
  github_document:
    toc: true
---

```{r}
library(tidyverse)
library(sf)
```

What I've tried to do here is first group incidents by country, then do an individual section for each event. You'll see that each section is headlined by the ISO 3-digit country code and the unique identifier I added to the Metabiota dataset in Google Sheets.

For general administrative boundaries, I'm using [the GADM database](https://gadm.org/). When I start a new country, I download that country's geopackage into the directory "./input_data/country/GADM/", where I extract the file to that directory instead of letting the unzipper create a new directory. 

Because these files are large, you'll want to use [Github's Large File Storage](https://git-lfs.github.com/) to track them. I would do the following commands:

* git lfs track "*.zip"
* git lfs track "*.gpkg"

As of now, my goal is to have each incident with the following fields:

* unique identifier
* polygon geometry, if used
* point geometry

We may want to consider adding a memo field to contain important information about each incident but for now I've not done that. 

What you might do is create a new R notebook and just start the document like below (i.e. ignore all the stuff above here). Then when you're done, we can just append it to this doc. 



## India

### ind_88

According to [this 2006 paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3373078/): “No definitive information about the possible index case exists.”

So, for now, let’s use the GADM layer to at least find the Siliguri area polygon.

First, let’s check the layers.

```{r}
st_layers("./input_data/India/gadm36_IND.gpkg")
```

So I think the best strategy for using this data is to start at the lowest level. Then you can also search across multiple geographies in a single table.

```{r}
india_gadm_3 <- st_read("./input_data/India/gadm36_IND.gpkg", layer="gadm36_IND_3") %>%
  janitor::clean_names()
```

Here we find what appears to be the correct polygon for Siliguri.

As you can see here, the select and mutate function calls are what create the end product mentioned in the intro section.

For precision, I've used "centroid" and "precise" --- with "precise" being the case when we are able to identify the exact location.

```{r}
ind_88 <- india_gadm_3 %>%
  filter(str_detect(name_3, "Siliguri")) %>%
  select(poly_geom = geom) %>% # this right here is where we start building the end dataset
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid") %>%
  mutate(reuters_id = 88,
         country_iso = "BGD")
```


```{r}
ind_88_point <- tribble(
  ~latitude, ~longitude,
  26.755302, 88.407176
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```




## Bangladesh

Get GADM data for repeated use. 

```{r}
st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")
```

```{r}
bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg", layer="gadm36_BGD_4") %>%
  janitor::clean_names()

```


### bgd_87

[This paper identifies an outbreak](https://onlinelibrary.wiley.com/doi/full/10.1002/rmv.2010) in Meherpur, Bangladesh from April-May 2001. 

"The first NiV outbreak was reported in April 2001 from a village in district Meherpur, Bangladesh with 13 confirmed cases and 9 (69.2%) deaths."

In Bangladesh, the administrative breakdown goes:

* Central government (level 0)
* Divisions (level 1)
* Districts, aka "zila" (level 2)
* Sub-districts, aka "upazila" (level 3)
* Union councils, municipalities and city corporations (level 4)

The union councils are rural areas and villages fall beneath those. For now, we don't have a good villages dataset. 

The above paper leads us to [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf), which says:

"Surveys took place in the villages of Chandpur (population 604), where persons who died or were hospitalized had resided..."

Walsh, M. G. (2015) place this location at 23.776652, 88.645198. Let's check if we can be more accurate. 

First let's create the point. 

```{r}
bgd_87_point <- tribble(
  ~latitude, ~longitude,
  23.776652, 88.645198
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```


Now, we can turn to [this official small-area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Meherpur.pdf) to find the smallest geography containing the village. See page 14. 

```{r}
bgd_87 <- bgd_gadm_4 %>%
  filter(name_2 == "Meherpur") %>%
  filter(name_4 == "Pirojpur")
```

Now let's compare the Walsh point to our current geography.

```{r}
ggplot() + 
  geom_sf(data=bgd_87) + 
  geom_sf(data=bgd_87_point, color="red")
```
We can see this is an improvement in geography. We'll use our new polygon. 

```{r}
bgd_87 <- bgd_gadm_4 %>%
  filter(name_2 == "Meherpur") %>%
  filter(name_4 == "Pirojpur") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid")  
```



### bgd_76

This outbreak in Naogoan is also profiled in [the same paper used for bgd_87]((https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3323384/pdf/04-0701.pdf)), the outbreak in Chandpur village, Meherpur district. 

Unfortunately, the paper never tells us where the actual index case occured. Here's two relevant passages: 

"Surveys were conducted in the adjacent villages of East Chalksita (population 529) and Biljoania (population 481); suspected deaths and hospitalizations caused by Nipah virus infection were reported from both villages."

and 

"The index case occurred in a 12-year-old boy with symptom onset on January 11, 2003, and the last case occurred in 12-year-old girl on January 28 (Figure B)."

But they never actually tell us where the index case lived.  

In this case, we're once again left with a large area and unable to narrow it down. This needs additional reporting. 


In [this report from ICDDR,B](http://dspace.icddrb.org/jspui/bitstream/123456789/4811/1/2003-ICDDRBHealthScienceBulletin-Vol1%285%29-English.pdf), we learn:

"Between 11 and 28 January 2003, another outbreak of severe illness including
features of encephalitis was reported affecting at least 17 residents (range 4-
42 years) of Chalksita and Biljoania villages (located 45 km north-east of
Rajshahi) in Naogaon district; 8 people died."


Walsh used 24.814569, 88.927434, but that appears to be a centroid. 


```{r}
bgd_76_point <- tribble(
  ~latitude, ~longitude,
  24.807898, 88.638664
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```

On page 22 of [this small-area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Naogaon.pdf), we can see in the index that Chak Sita and Bil Joania are in Sreemantapur Union (in NIAMATPUR UPAZILA).

We can use that information to find the level 4 polygon in the GADM data.

FYI: The GADM dataset spells it Naogaon. 

```{r}
bgd_76 <- bgd_gadm_4 %>%
  filter(name_2 == "Naogaon") %>%
  filter(name_4 == "Sreemantapur")
```


Now let's visualize the Walsh point with our level 4 polygon. 

```{r}
ggplot() + 
  geom_sf(data=bgd_76) +
  geom_sf(data=bgd_76_point, color="red")
```






### bgd_73

http://dspace.icddrb.org/jspui/bitstream/123456789/4816/1/2004-ICDDRBHealthScienceBulletin-Vol2%282%29-English.pdf



### bgd_72

This is an outbreak in Rajbari District, Bangladesh during January and February 2004. It occurred at the same time as another outbreak in Manikganj District, Bangladesh, which isn't in the Metabiota data, so we'll probably need to add it. I have labeled it as record x7 in the potential adds tab of the spreadsheet. 

According to this paper:

"On 17 January 2004, a 9-year-old child from Kazipara village was brought to the emergency room of Goalanda health complex in Rajbari district, Bangladesh. The patient had a three-day history of fever and worsening neurological signs: he died within 15 minutes of admission. Between 17 and 21 January, 6 additional children (aged between 2 and 15 years) and a 28 year-old woman (mother of the youngest child) also died with the same symptoms. All patients came from the neighbouring Kazipara and Juran Molla Para villages and lived in two groups of houses located within 1 km of each other."

And we learn from [this 2008 paper](https://academic.oup.com/cid/article/46/7/977/292786) that it's in Goalanda subdistrict. 

So first, let's use GADM to try to identify the polygon for Goalanda subdistrict. GADM appears to spell Goalanda as Goalandaghat.


```{r}
bgd_72 <- bgd_gadm_4 %>%
  filter(name_3 == "Goalandaghat") %>%
  select(name_3) %>%
  group_by(name_3) %>%
  summarise(geom = st_union(geom)) %>%
  select(poly_geom = geom)
```

Walsh placed Nipah at the following location: 23.748507, 89.655376. Let's see how it compared to the GADM polygon.

```{r}
bgd_72_point <- tribble(
  ~latitude, ~longitude,
  23.748507, 89.655376
) %>%
  st_as_sf(coords=c("longitude", "latitude"), crs=4326) 
```

Let's put 'em on a map.

```{r}
ggplot() +
  geom_sf(data=bgd_72) +
  geom_sf(data=bgd_72_point)
```
How far from the polygon to the point?

```{r}
bgd_72_8858 <- bgd_72 %>%
  st_transform(8858)

bgd_72_point_8858 <- bgd_72_point %>%
  st_transform(8858)

st_distance(bgd_72_point_8858, bgd_72_8858)
```
So there's about a half-kilometer distance between the point used by Walsh and the polygon in which Gurley said the outbreak occurred. 

If we create a centroid, how far between Walsh and our centroid?

```{r}
bgd_72_8858_centroid <- bgd_72 %>%
  st_transform(8858) %>%
  st_centroid()

st_distance(bgd_72_point_8858, bgd_72_8858_centroid)

```
So now there's about a 1km difference.

We should try to nail down the location of Kazipara village.

By using the [Rajbari small area atlas](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Rajbari.pdf), we're able to figure out that Kazi Para, as its spelled in the document, is in GOALANDAGHAT PAURASHAVA. 

If we go back to our map, we can see that is a level 4 geography. While we don't know the exact village location, we've now narrowed it down to a tiny area. The Goalandaghat Paurashava is in red. 

```{r}
test <- bgd_gadm_4 %>%
  filter(name_3 == "Goalandaghat") %>%
  filter(name_4 == "Goalandaghat Paurashava") 

  ggplot() + 
  geom_sf(data=bgd_72) +
  geom_sf(data=test, color="red")
```
So we're now going to use that as our polygon and take the centroid of that. 

```{r}
bgd_72 <- bgd_gadm_4 %>%
  filter(name_3 == "Goalandaghat") %>%
  filter(name_4 == "Goalandaghat Paurashava") %>%
  select(poly_geom = geom) %>%
  mutate(point_geom = st_centroid(poly_geom),
         precision = "centroid")
  
```

Absent the exact location of the village, this is very accurate. 



### bgd_68 

http://dspace.icddrb.org/jspui/bitstream/123456789/4852/1/2005-ICDDRBHealthScienceBulletin-Vol3%284%29English.pdf


