---
title: "Adding data - Allison's work"
output: word_document
---

Set working directory for Allison's Windows laptop

```{r}

setwd("C:/Users/0146156/Documents/mb_checks/reuters-news-data_disease_gis")

```

Load libraries

```{r}

library(tidyverse)
library(sf)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)

```

Get GADM data. 

```{r}

st_layers("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg")

bgd_gadm_2 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_2") %>%
              janitor::clean_names()

bgd_gadm_3 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer = "gadm36_BGD_3") %>%
              janitor::clean_names()

bgd_gadm_4 <- st_read("./input_data/Bangladesh/gadm/gadm36_BGD_gpkg/gadm36_BGD.gpkg",
                      layer="gadm36_BGD_4") %>%
              janitor::clean_names()

st_layers("./input_data/India/gadm36_IND.gpkg")

ind_gadm_3 <- st_read("./input_data/India/gadm36_IND.gpkg",
                      layer = "gadm36_IND_3") %>%
              janitor::clean_names()

```

Get other spatial data.

```{r}

st_layers("./input_data/India/Kerala")

ind_taluks <- st_read("./input_data/India/Kerala", layer = "taluk") %>%
  rename(geom = geometry, oid = OID_, taluk = TALUK, state = STATE, district = DISTRICT)

st_layers("C:/Users/0146156/Documents/mb_checks/stanford_level_3")

phl_level_3 <- st_read("C:/Users/0146156/Documents/mb_checks/stanford_level_3")

```

Setting up function that returns variables based on single GADM polygon.

```{r}

grab_polygon <-function(geom, id) {
  geom %>% select(poly_geom = geom) %>% 
           mutate(point_geom = st_centroid(poly_geom),
                  precision = "centroid",
                  reuters_id = id,
                  country_iso = "BGD")
}

```


#bgd_90

Viral sequence is listed as a primary case in a male patient in April 2018 Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Rajshahi, no upazilla mentioned. 

*** Coding to district

```{r}

bgd_90 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Rajshahi")) %>%
  grab_polygon(90)

```

#bgd_91

Viral sequence is listed as a primary case in a 26 year old male patient in February 2018 in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Faridpur, no upazilla mentioned.

*** Coding to Faridpur district

```{r}

bgd_91 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Faridpur")) %>%
  grab_polygon(91)

```

#bgd_92

Viral sequence is listed as a primary case in a 13 year old male in 2017 in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). Location given is Uttar Raghobpur in Pabna district.

Uttar Raghobpur is not an upazilla or a union. [This website](https://www.thebangladesh.net/rajshahi-division/pabna-district/pabna-sadar-upazila-pabna/malanchi-union-pabna/uttar-raghabpur-mouza-pabna-sadar-pabna.html) suggests it is a lower administrative division called a 'mouza'. The site says it is in Malanchi union of Pabna Sadar, which is in GADM.

The small area atlas [here](http://203.112.218.65:8008/WebTestApplication/userfiles/Image/Atlas/Pabna.pdf) confirms on page 40 that Uttar Raghabpur is part of Malanchi union.

*** Coding to Malanchi for now, but could later code to the mouza 

```{r}

bgd_92 <- bgd_gadm_4 %>% filter(str_detect(name_2, "Pabna")) %>%
  filter(str_detect(name_4, "Malanchi")) %>%
  grab_polygon(92)

```
#bgd_93

Viral sequence is listed as a primary case in a 12 year old male in 2017 in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). Location given is Uttar Raghobpur in Pabna district. Location is apparently Joykalidangi, Faridpur, but "Joykalidangi" does not exist as written outside this academic paper.

The given date is February 15, 2017. IEDCR's site says they investigated an encephalities cluster in Faridpur in late January, see [here](https://old.iedcr.gov.bd/website/index.php/outbreak/203-outbreak2017)

```{r}

bgd_gadm_4 %>% filter(str_detect(name_4, "Dangi"))

```

There is a Dangi union in Faridpur. 

No information about this outbreak in other sources.

*** Not coding for now.


#bgd_97

IEDCR narrative lists early outbreaks: "In the last 10 years we have identified another eight Nipah outbreaks, all occurring between December and May. The Nipah outbreaks have been identified in Noagoan (2003), Rajbari (2004), Faridpur (2004), Tangail (2005), Thakurgaon (2007) Kushtia (2007), Manikgonj and Rajbari ( 2008), Faridpur ( 2010)." [source](https://webcache.googleusercontent.com/search?q=cache:NZKNrjYhTzgJ:iedcr.gov.bd/pdf/files/nipah/Nipah-for-web.pdf+&cd=1&hl=en&ct=clnk&gl=ca)

IEDCR outbreak investigation list includes an April 2007 Nipah outbreak in Kushtia [here](https://old.iedcr.gov.bd/website/index.php/outbreak/85-outbreak-list-of-2007)

Deka, 2018 includes this outbreak and codes it to the polygon for Sadar upazilla, Kushtia, citing Walsh.

Walsh in turn cites two sources:
1. ProMED 20050125.0267. Undiagnosed disease-Bangladesh (Tangail)(02): Nipah confirmed. 
2. International Society for Infectious Diseases; 2005.
Bangladesh Institute of Epidemiology Disease Control and Research. Nipah Virus Outbreak from Date Palm Juice. Heal Sci Bull 2005;3:1–5.

As both of these papers predate the outbreak (!) it seems more likely that the table should have referred to this reference:
Homaira N, Rahman M, Hossain MJ et al. Cluster of Nipah virus
infection, Kushtia District, Bangladesh, 2007. PLoS One
2010;5:e13570.

The paper says the outbreak was in a village within the sadar upazilla, and more specifically that: "The grave of a Muslim religious leader, who died in 2001, is in the outbreak village and his followers from different villages gather annually for an anniversary ceremony. On the third day of the
index case’s illness, on March 19, 2007, this religious ceremony
was in the courtyard adjacent to the grave of the religious leader."

ProMED bulletin from April 12, 2007: "According to the hospital and Civil Surgeon's office, the disease 1st broke out in Khajanagar and Kaburhat villages in Kushtia district a week ago."

* Coding to Sadar upazilla

```{r}

bgd_97 <- bgd_gadm_3 %>% filter(str_detect(name_2, "Kushtia")) %>%
  filter(str_detect(name_3, "Sadar")) %>%
  grab_polygon(97)

```


#bgd_98

This adds a second location to the incident first described in bgd_41, see notes there.

```{r}

bgd_98 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Khetlal")) %>%
  grab_polygon(98)
  
```


#phl_99

This adds a second location to the incident described in phl_22, see notes there. 

```{r}

phl_99 <- phl_level_3 %>% filter(str_detect(name_3, "Midtungok")) %>%
  rename(geom = geometry) %>%
  grab_polygon(99) %>% 
  mutate(country_iso = "PHL")

```

#bgd_100

Walsh, 2015 lists a 2014 case in Naogaon district, Manda upazilla, but the reference is to an unrelated paper, so something has been lost in the data processing. 

Naogaon district appears in IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

Walsh puts that case at February 2014, but perhaps that is because the IEDCR bulletin was released in that month and appeard in ProMED etc then? 

There are local media reports of an unconfirmed case in Naogaon, published on and around January 12, 2014. See for example the Dhaka Tribune: "Sources said it was also suspected that the deaths of four more people [...] may also be linked to NiV infection. The deceased [include] Mojibur Rahman, 20, from Naogaon [...] Doctors at the RMCH said they suspected that the deaths may have been caused by Nipah infection as all four patients had the history of drinking raw date-palm sap."

Story is archived online (here)[https://web.archive.org/web/20140424013844/http://www.dhakatribune.com/safety/2014/jan/12/fears-nipah-virus-outbreak-following-2-confirmed-deaths]

Dhaka Tribune said lab confirmation was not available and in fact one of the four (not identified) had tested negative: "Prof Dr Mahmudur Rahman, director of the IEDCR, however, told the Dhaka Tribune yesterday [11 Jan 2014] that the Nipah virus was not found in the single sample of the deceased that had so far been tested at the lab. Samples from 2 other deceased would be tested by Monday [13 Jan 2014], he said, adding that sample from another patient had not yet arrived at the IEDCR." [here](https://promedmail.org/promed-post/?id=2181682)

Financial Express, via ProMED pinpoints the village: "According to hospital sources, the deceased were identified as [...] an adult from Naogoan's Tentulia village" [here](https://promedmail.org/promed-post/?id=2168940)

Tentulia is a Union in Manda, the upazilla named in Walsh, 2015. 

*** Given that consistency, coding to Tentulia.

```{r}

bgd_100 <- bgd_gadm_4 %>% filter(str_detect(name_3, "Manda")) %>%
  filter(str_detect(name_4, "Tentulia")) %>%
  grab_polygon()

```


#bgd_101 / bgd_102 / bgd_103 / bgd_104

All of these 2012 cases are drawn from an archived WHO table which has been incorporated into for example Deka (CDC's) data but not Metabiota's database. One academic paper strongly suggests they are all connected to the well documented 2012 Joypurhat outbreak but it is not written by local experts and may be a simple misinterpretation of the WHO data: "The final outbreak was in 2012 in Joypurhat, Rajshshi, Natore, and Rajbari with a fatality rate of 83% (n=12)." (paper here)[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6415930/]

Sought more details from IECDR website, ProMED, Daily STar, BDNews, FluTrackers, Factiva and general Google searches.

Rajshahi:

A little concerned this is a coding error because one of the Joypurhat cases apparently died at a hospital in Rajshahi. See ProMED: "Of the 5 deaths, 2 are from Khetlal and the others from the municipality. They died in Ziaur Rahman Medical College Hospital in Bogra and Rajshahi Medical College Hospital. On 21 Jan [2012], a 5-member IECDR expert team went to Bogra and Rajshahi to collect blood and other samples." (source)[https://promedmail.org/promed-post/?id=1022056]

There is a slightly better document Rajshani case for 2013, which we coded under bgd_34. 

Then again it is clearly included in the WHO's reporting table, archived online (here)[http://web.archive.org/web/20190922101016/http://www.searo.who.int/entity/emerging_diseases/links/nipah_virus_outbreaks_sear/en/]

Natore, Rajbari, Gopalganj: 

No indication of a case there in other sources.

*** Not coding these

#bgd_105

This is a potential cause in Pataukhali district, pulled from an archived news story. After discussion with Deb, excluding because time from sap drinking to death seems too short.

"Source: The Daily Star, Staff Correspondent
Date: 10 February 2012
Nipah virus claimed the life of an SSC examinee at Patuakhali General Hospital yesterday. The victim,
Nargis Akhter, 16, daughter of Hasan Sikder of Gabua village under Patuakhali Sadar upazila, was
appearing in the ongoing SSC examinee from Gabua Janata High School. Family and hospital sources said
Nargis fell sick after having date juice at around 8:00am. She was taken to the hospital where she died at
around 1:30pm."
[Archived story](https://web.archive.org/web/20120412182445/http://saarc-sdmc.nic.in/pdf/saarc_news/news227.pdf)

#bgd_106

Viral sequence is listed as a primary case in a 7 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Kurigram and subdistrict Fulbari. 

IECDR's list of 2013 cases lists a 7 year old male who died in Kurigram that year.

*** Coding to Fulbari upazila, which in GADM is transliterated as 'Phulbari'

```{r}

bgd_106 <- bgd_gadm_3 %>% filter(str_detect(name_2, "Kurigram")) %>%
  filter(str_detect(name_3, "Phulbari")) %>%
  grab_polygon(106)

```
#bgd_107

Viral sequence is listed as a primary case in a 7 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Gaibandha and subdistrict Polashbari. 

IEDCR's list of 2013 cases includes a 9 year old male who died in Gaibandha that year.

*** Coding to Palashbari upazilla 

```{r}

bgd_107 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Palashbari")) %>%
  grab_polygon(107)

```

#bgd_108

Viral sequence is listed as a primary case in a 43 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Pabna and subdistrict is Chatmohor. 

IEDCR's list of 2013 cases includes a 43 year old male who died in Pabna that year, and a three year old in the same district who survived. 

From BDNews24 via ProMED: "The new deaths were from Pabna and Rajshahi while one 8-month old boy is struggling for life in the Rajshahi Medical College Hospital." [link](https://promedmail.org/promed-post/?id=1518442)

*** Coding to Chatmohar upazilla

```{r}

bgd_108 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Chatmohar")) %>%
  grab_polygon(108)

```

#bgd_109

Viral sequence is listed as a primary case in a 48 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Nilphamari and subdistrict is Sayedpur. 

IEDCR's list of 2013 cases includes a 48 year old male who died in Nilphamari that year.

*** Coding to Sayedpur upazilla, transliterated in GADM to Saidpur

```{r}

bgd_109 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Saidpur")) %>%
  grab_polygon(109)

```

#bgd_110

Viral sequence is listed as a primary case in a 22 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Rangpur and subdistrict is Sadar.

IEDCR's list of 2014 case counts by district identifies four in Rangpur.

*** Coding to the Sadar upazilla of Rangpur.

```{r}

bgd_110 <- bgd_gadm_3 %>% filter(str_detect(name_2, "Rangpur")) %>%
filter(str_detect(name_3, "Sadar")) %>%
  grab_polygon(110)

```
#bgd_111

Viral sequence is listed as a primary case in a 22 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Faridpur and subdistrict is Saltha.

IEDCR's list of 2014 case counts by district identifies five in Faridpur

*** Coding to Saltha upazilla

```{r}

bgd_111 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Saltha")) %>%
  grab_polygon(111)

```

#bgd_112 / bgd_113

From Dhaka Tribune, archived through the Wayback Machine: "Health specialists have expressed concerns about a potential fresh outbreak of the deadly Nipah virus (NiV), following the confirmation of two recent deaths from NiV infection and four more suspected fatalities.

"Laboratory tests at the Institute of Epidemiology, Disease Control and Research (IEDCR) has confirmed that two boys – aged 13 and 14 – died from NiV infection at Shibaloy and Ghior of Manikganj district on the last week of December last year.Both of them died after being infected with the virus by drinking raw date-palm sap (Khejurer Rosh)."

[link](https://web.archive.org/web/20140525203226/http://www.dhakatribune.com/safety/2014/jan/12/fears-nipah-virus-outbreak-following-2-confirmed-deaths)

*** Coding to Shibalaya and Ghior upazillas

```{r}

bgd_112 <- bgd_gadm_3 %>% 
  filter(str_detect(name_3, "Shibalaya")) %>%
  grab_polygon(112)

bgd_113 <- bgd_gadm_3 %>% 
  filter(str_detect(name_3, "Ghior")) %>%
  grab_polygon(113)

```


#bgd_114

Viral sequence is listed as a primary case in a 14 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Magura and subdistrict Sadar.

IEDCR's list of 2014 case counts by district identifies five in Magura

** Coding to Magura Sadar upazilla

```{r}

bgd_114 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Magura Sadar")) %>%
  grab_polygon(114)

```
#bgd_115 

Viral sequence is listed as a primary case in a 7 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Gopalganj and subdistrict is Mokshedpur.

GADM data transliterates as more standard Muksudpur, which data [here](https://moa.portal.gov.bd/sites/default/files/files/moa.portal.gov.bd/page/7bd6781d_6460_4531_8df7_84c90ed2c8b4/Gopalganj-12.03.2020.pdf) makes clear is equivalent. 

*** Coding to Muksudpur upazilla

```{r}

bgd_115 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Muksudpur")) %>%
  grab_polygon(115)

```

#bgd_116

Viral sequence is listed as a primary case in a 36 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Madaripur and subdistrict Shibchar.

*** Coding to upazilla Shib Char 

```{r}

bgd_116 <- bgd_gadm_3 %>% filter(str_detect(name_3, "Shib Char")) %>%
  grab_polygon(116)

```

#bgd_117

Viral sequence is listed as a primary case in a 42 year old male in Rahman et al 2021, [here](https://www.sciencedirect.com/science/article/pii/S1201971220322451#tbl0020). District is Rajshahi and subdistrict Bagha.

***Coding to Bagha upazilla

```{r}

bgd_117 <- bgd_gadm_3 %>% filter(str_detect(name_2, "Rajshahi")) %>%
  filter(str_detect(name_3, "Bagha")) %>%
  grab_polygon(117)

```

#bgd_118

Mymensingh district appears on the IEDCR case list for 2013 [here](https://webcache.googleusercontent.com/search?q=cache:bjtu4Hc2QRQJ:www.iedcr.gov.bd/index.php/component/content/article/11/135-23-rd-february-2013-nipah-outbreak+&cd=1&hl=en&ct=clnk&gl=ca), and is repeated in an academic paper that cites IEDCR for districts with cases that year [here](http://ebaub.edu.bd/journal/ej/files/vol3/jc3_6.pdf).

[This](https://www.thenewhumanitarian.org/news/2013/02/05/fresh-outbreak-nipah-virus-bangladesh) news story notes that the Dhaka cases coded earlier, bgd_24, came from sap coillected in from Mymensingh. But there are several cases in teh bulletin, suggesting another spillover

*** Coding to Mymensingh district

```{r}

bgd_118 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Mymensingh")) %>%
  grab_polygon(118)


```

#bgd_119

Nilphamari district appears on the IEDCR case list for 2013 [here](https://webcache.googleusercontent.com/search?q=cache:bjtu4Hc2QRQJ:www.iedcr.gov.bd/index.php/component/content/article/11/135-23-rd-february-2013-nipah-outbreak+&cd=1&hl=en&ct=clnk&gl=ca), and is repeated in an academic paper that cites IEDCR for districts with cases that year [here](http://ebaub.edu.bd/journal/ej/files/vol3/jc3_6.pdf).

In #bgd_109 we coded a 48 year old who appears in the 2013 list and was sequenced. But IEDCR lists two other cases in the district later in the table, in a 55-year-old male and 45-year-old female, so coding this as a separate spillover.

*** Coding to Nilphamari

```{r}

bgd_119 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Nilphamari")) %>%
  grab_polygon(119)

```

#bgd_120

Naogaon district appears on the IEDCR case list for 2013 [here](https://webcache.googleusercontent.com/search?q=cache:bjtu4Hc2QRQJ:www.iedcr.gov.bd/index.php/component/content/article/11/135-23-rd-february-2013-nipah-outbreak+&cd=1&hl=en&ct=clnk&gl=ca).

Coding this separately from bgd_32, a specific spillover involving an 8 year old in the same district in 2013, because IEDCR lists additional patients later in thier table. 

*** Coding to Naogaon

```{r}

bgd_120 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Naogaon")) %>%
  grab_polygon(120)

```

#bgd_121

Rajbari district appears on the IEDCR case list for 2013 [here](https://webcache.googleusercontent.com/search?q=cache:bjtu4Hc2QRQJ:www.iedcr.gov.bd/index.php/component/content/article/11/135-23-rd-february-2013-nipah-outbreak+&cd=1&hl=en&ct=clnk&gl=ca) as a 24 year old male who died.

* Coding to Rajbari district

```{r}

bgd_121 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Rajbari")) %>%
  grab_polygon(121)

```

#bgd_122

Jhenaidah district appears on the IEDCR case list for 2013 [here](https://webcache.googleusercontent.com/search?q=cache:bjtu4Hc2QRQJ:www.iedcr.gov.bd/index.php/component/content/article/11/135-23-rd-february-2013-nipah-outbreak+&cd=1&hl=en&ct=clnk&gl=ca) as a 11 year old male who died.

* Coding to Jhenaidah district

```{r}

bgd_122 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Jhenaidah")) %>%
  grab_polygon(122)

```

#bgd_123

Faridpur district appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

* Coding to Faridpur district

```{r}

bgd_123 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Faridpur")) %>%
  grab_polygon(123)

```

#bgd_124

Shariatpur district appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

* Coding to Shariatpur district

```{r}

bgd_124 <- bgd_gadm_2 %>%
  filter(str_detect(name_2, "Shariatpur")) %>%
  grab_polygon(124)

```

#bgd_125

Kushtia district appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

* Coding to Kushtia district

```{r}

bgd_125 <- bgd_gadm_2 %>%
  filter(str_detect(name_2, "Kushtia")) %>%
  grab_polygon(125)

```

#bgd_126

Rajshahi district appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

Dhaka Tribune reported on January 12: "Sources said it was also suspected that the deaths of four more people including a five-year-old boy at the Rajshahi Medical College Hospital (RMCH) last week, may also be linked to NiV infection. The deceased were Sayedur Rahman, 45, from Bagha upazila of Rajshahi; Naznur Rahman, 5, from Pabna’s Ishwardi; Mojibur Rahman, 20, from Naogaon; and Hasanuzzaman, 20, from Natore’s Bagatipara. Doctors at the RMCH said they suspected that the deaths may have been caused by Nipah infection as all four patients had the history of drinking raw date-palm sap."

Story is archived online (here)[https://web.archive.org/web/20140424013844/http://www.dhakatribune.com/safety/2014/jan/12/fears-nipah-virus-outbreak-following-2-confirmed-deaths]

* Could code to Bagha upazila?
*** Not sure what to do with this 

#bgd_127

Natore district appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

Dhaka Tribune reported on January 12: "Sources said it was also suspected that the deaths of four more people including a five-year-old boy at the Rajshahi Medical College Hospital (RMCH) last week, may also be linked to NiV infection. The deceased were Sayedur Rahman, 45, from Bagha upazila of Rajshahi; Naznur Rahman, 5, from Pabna’s Ishwardi; Mojibur Rahman, 20, from Naogaon; and Hasanuzzaman, 20, from Natore’s Bagatipara. Doctors at the RMCH said they suspected that the deaths may have been caused by Nipah infection as all four patients had the history of drinking raw date-palm sap."

Story is archived online (here)[https://web.archive.org/web/20140424013844/http://www.dhakatribune.com/safety/2014/jan/12/fears-nipah-virus-outbreak-following-2-confirmed-deaths]

Financial Express, via ProMED: "According to hospital sources, the deceased were identified as a 45 year old adult and a 5 year old child of Ishwardi's Bahbani village, [an adult] from Naogoan's Tentulia village and a 20 year old adult from Natore's Bagatipara area [all in Rajshahi Division]." [here](https://promedmail.org/promed-post/?id=2168940)

More from Dhaka Tribune the same day, via ProMED: "The deceased were a man of 45 from Bagha upazila of Rajshahi; a 5 year old from Pabna's Ishwardi; an individual of 20 from Naogaon; and another 20 year old from Natore's Bagatipara [all in Rajshahi division].

Doctors at the RMCH said they suspected that the deaths may have been caused by Nipah infection, as all 4 patients had the history of drinking raw date-palm sap.

Prof Dr Mahmudur Rahman, director of the IEDCR, however, told the Dhaka Tribune yesterday [11 Jan 2014] that the Nipah virus was not found in the single sample of the deceased that had so far been tested at the lab. Samples from 2 other deceased would be tested by Monday [13 Jan 2014], he said, adding that sample from another patient had not yet arrived at the IEDCR." [here](https://promedmail.org/promed-post/?id=2181682)


* Could code to Bagatipara
* Look in other sources
*** Come back to this


#bgd_128

Dinajpur district appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

* Code to Dinajpur district

```{r}

bgd_128 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Dinajpur")) %>%
  grab_polygon(128)

```
#bgd_129

Chapi Nawabganj/Chapainawabganj appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

* Code to Chapi Nawabganj district

This district name does not appear in GADM, but based on the associated upazillas, it appears under the name "Nawabganj". 

```{r}

bgd_129 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Nawabganj")) %>%
  grab_polygon(129)

```

#bgd_130

Madaripur appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

* Code to Madarirpur district

```{r}

bgd_130 <- bgd_gadm_2 %>% 
  filter(str_detect(name_2, "Madaripur")) %>%
  grab_polygon(130)

```

#bgd_131

Dhaka appears on the IEDCR case list for Dec 2013 - February 27, 2014, archived through Google's cache [here](https://webcache.googleusercontent.com/search?q=cache:lL29-kGy1uEJ:https://www.iedcr.gov.bd/pdf/files/nipah/Nipah%2520Infection_27.02.2014.pdf+&cd=3&hl=en&ct=clnk&gl=ca)

Dhaka Tribune describes cases

* Coding to Dhaka district

```{r}

bgd_131 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Dhaka")) %>%
  grab_polygon(131)

```

#bgd_132

Local media report from January 17, 2020: "The patient was identified as Rashida, 20, daughter of Md Babul Hossain of Gopalganj district." [link](https://en.prothomalo.com/bangladesh/Nipah-case-confirmed-in-Khulna)

Confirmation from Dhaka Tribune: "“A medical board has confirmed her infection by Nipah virus. As Rashida’s infection is a risk to other patients, she is being treated separately at the hospital’s Medicine unit 1,” said SM Kamal Hossain, chief of KMCH Medicine Department." [link](https://archive.dhakatribune.com/bangladesh/nation/2020/01/17/fresh-nipah-infected-patient-at-kmch)

* Coding to Gopalganj district

```{r}

bgd_132 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Gopalganj")) %>%
  grab_polygon(132)

```

#bgd_133

Coding the 2011 Dinajpur cluster identified in [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4675679/). In the Metabiota data it was included in bgd_44 but we are breaking out separate clusters.

The paper posits an independent spillover in Dinajpur: "Of the five cases in the Dinajpur cluster, two probable cases had drunk raw date palm sap prior to illness onset. Two cases, one confirmed and one probable, came in contact with these probable cases."

The paper gives a date range from Dec 2010 to 2011, but The Daily Star, archived by ProMED, said in Feb 2011 that people had died in Dinajpur from Nipah since January 31, 2011, see [here](https://promedmail.org/promed-post/?id=653198) and [here](https://web.archive.org/web/20110214192935/http://www.thedailystar.net/newDesign/latest_news.php?nid=28442).

The Daiy Star reported on Feb 11, 2011 that someone had died of Nipah in Dinajpur: "Masuma Akhter Moni, 11, of Belaichandi village in Parbatipur upazila of Dinajpur [...] Moni was admitted to the hospital with high fever on Tuesday."

According to a FluTrackers thread the paper later reported this case: "Monoranjan, 4, son of Dhirendra Nath of Kaharol upazila of Dinajpur, was admitted to the RMCH with high fever on Friday. He died at about 9:00am Saturday, reports our Rangpur correspondent."

And on February 14: "Nipah encephalitis claimed two more lives at Rangpur medical College Hospital (RMCH) on Monday, raising the death toll to 33. Anis, 8, son of Alam Miah of Atowari, Dinajpur died at about 2:30am" [source](https://web.archive.org/web/20110215145629/http://www.thedailystar.net/newDesign/latest_news.php?nid=28463)

FluTrackers quotes another local media story taht is no longer online on Feb 24: "Maloti Rani, 35, wife of Manoranjan of village Shahapur under Kaharol upazila in Dinajpur, was admitted at 11am" [here](https://flutrackers.com/forum/forum/emerging-diseases-other-health-threats-alphabetical-i-thru-z/nipah-virus/100720-bangladesh-36-dead-from-nipah-virus/page7)

Even with all that, it's not clear which upazilla the initial spillover was in, so coding to Dinajpur.

```{r}
bgd_133 <- bgd_gdam_2 %>% 
  filter(str_detect(name_2, "Dinajpur")) %>%
  grab_polygon(133)

```
#bgd_134

A Rangpur district cluster is described in [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4675679/). It does not appear in the Metabiota data. Media reports were mixed in with the lage outbreak in Lalmonirhat around the same time but the paper describes them as separate spillovers: "Of the eight cases in the Rangpur cluster, proxies of three probable cases reported that the cases regularly drank fermented date palm sap, a traditional liquor known locally as tari, from a common source. They consumed the drink daily during the 30 days prior to their illness onset. These probable cases did not have any history of drinking raw sap or of having contact with a Nipah case. Three confirmed cases became ill following close contact with one of these probable cases. Two other cases from this cluster, one confirmed and one probable, had no reported history of consuming raw or fermented date palm sap or contact with a Nipah case."

Can we make this location any more precise? 

The Daily Star via ProMED reported on Feb 3, 2011: "The dead were identified as [individuals whose ages and locations are], 10 and 40, of South Goddimari village; 23 of West Bejpur; 50 of Tongbhanga; and another 8, of local bus terminal area in Hatibandha
upazila [sub-district] of Lalmonirhat." [source](https://promedmail.org/promed-post/?id=644637)

The DAily Star via ProMED reported on Feb 4, 2011: "The government on Friday [4 Feb 2011] confirmed the disease as Nipah
virus encephalitis that claimed at least 21 people over the last few days in Lalmonirhat and Rangpur."

Skipping detailed notes on villages with cases, but there is an upazilla mention: "The residents of Hatibandha and Gangachara upazilas have been terrified with the outbreak of the virus and many of them left their houses." - Gangachara is in Rangpur [source](https://promedmail.org/promed-post/?id=648343)

* Coding to Rangpur district based on academic paper

```{r}

bgd_134 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Rangpur")) %>%
  grab_polygon(134)

```
#bgd_135

This district was listed in Metabiota's data under bgd_44 in 2011; breaking it out here. 

The spillover is described in [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4675679/)

"The case reported by a private hospital in Dhaka, was a patient from Comilla District, in the eastern part of the country, who was admitted to that hospital with encephalitis. Her physicians suspected NiV as a possible cause of her illness since she had consumed raw date palm sap and sent her sera to IEDCR for testing. She was confirmed as a case of Nipah encephalitis. She drank raw date palm sap harvested from a tree in the courtyard of her house within 2 weeks prior to her illness onset, and other members from her family and neighbours also drank from the same pot during that time. However, no one else who had shared the sap with her reported any illness. Five of the six isolated cases reported a history of consuming raw date palm sap. None of the isolated cases had contact with any encephalitis cases prior to illness onset."

* Coding to district 

```{r}

bgd_135 <- bgd_gadm_2 %>% filter(str_detect(name_2, "Comilla")) %>%
  grab_polygon(135)

```

#bgd_136

Adding a small 2011 cluster in Rajbari described in [this paper](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4675679/)

"Both cases in the Rajbari cluster had drunk raw date palm sap within 3 weeks prior to their illness onset, but the second case (daughter of the first case) also came in contact with her mother after her onset of illness, which was 4 days prior to her onset of illness."

The Daily Star reported on Feb 11, 2011: "Meantime, three people from Rajbari were admitted to Dhaka Medical College Hospital last night with symptoms of Nipah virus, DMCH sources said.
The three are Abul Kalam, 55, his son Rusel, 18, and his nephew Abul Kalam, 18. Kalam's wife Asma Begum told this correspondent that five members of her family fell sick after taking some date juice on Monday. Two of them, however, recovered by taking some medicines at home." [source](https://www.thedailystar.net/news-detail-173755)

Not clear this is the same case.

Dhaka Mirror, via ProMED, reported on Jan 24, 2011: "The director of the Institute of Epidemiology, Disease Control and Research [IEDCR], Mahmudur Rahman, on Sunday [23 Jan 2011] said that 5 people were infected with the Nipah virus in Faridpur and Rajbari districts and 4 of them died this month [January 2011]."

